---
title: "Drug_dataset_V1"
author: "Aine Duffy"
date: "6/25/2020"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


  
```{r, include=F}

library(data.table)
library(stringr)
library(xlsx)
library(ggplot2) 
library(dplyr)
library(kableExtra)
library(DT)
library(scales)
library(gridExtra)
library(lme4)
```

### Overview of dataset 
Drug dataset:
- target genes
- sider V2 se matched to phecode
- se adr score
- omim matched to phecode
- coloc tissue and phenotype matched to phecode at pph4 >08
- Tau and OE score
``` {r, echo=F}
dataset=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data/Dataset/drug_sider2_omim_tau_coloc.txt',data.table=F)
dim(dataset) #7293455
head(dataset)
 dataset$entrez_id=as.character(dataset$entrez_id)
paste(length(unique(dataset$drugname)),'unique drugs,',length(unique(dataset$entrez_id)),'unique genes,',length(unique(dataset$PHECODE_SE)),'unique se,',length(unique(dataset$omim_phecode)),'unique omim phecode,',length(unique(dataset$coloc08_phecode)),'unique coloc2 phecode')
 "589 unique drugs, 658 unique genes, 672 unique se, 97 unique omim phecode, 765 unique coloc2 phecode"
phecodesource=dataset %>% distinct(coloc08_phecode,Source)
print(table(phecodesource$Source))
#both refers to the triplet being found in coloc results from saige and neale not just the phecode. (Why doesnt add to 765)
#693  saige     233 ukbb        94 both
uniquese=data.frame(phecode=unique(dataset$PHECODE_SE))
uniqueomim=data.frame(phecode=unique(dataset$omim_phecode))
matchseomim=intersect(uniqueomim, uniquese)
paste(length(matchseomim$phecode), 'phecodes are in unique omim phecodes and se')
#51 shared phecodes between omim and se
uniquecoloc=data.frame(phecode=unique(dataset$coloc08_phecode))
matchsecoloc=intersect(uniquecoloc, uniquese)
paste(length(matchsecoloc$phecode), 'phecodes are in unique coloc phecodes and se') 
"336 phecodes are in unique coloc phecodes and se"

```

### Types of phecodes
```{r fig.height=12, fig.width=14, echo=F}
phecodef=read.csv('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/phecode_definitions1.2.csv')
uniquecoloc$category=phecodef$category[match(uniquecoloc$phecode, phecodef$phecode)]
uniquecoloc$phenotypename=phecodef$phenotype[match(uniquecoloc$phecode, phecodef$phecode)]
uniquese$category=phecodef$category[match(uniquese$phecode, phecodef$phecode)]
uniquese$phenotypename=phecodef$phenotype[match(uniquese$phecode, phecodef$phecode)]

uniqueomim$category=phecodef$category[match(uniqueomim$phecode, phecodef$phecode)]
uniqueomim$phenotypename=phecodef$phenotype[match(uniqueomim$phecode, phecodef$phecode)]

uniquecoloc$source='coloc'
uniqueomim$source='omim'
uniquese$source='se'
allphecode=rbind(uniquese,uniquecoloc,uniqueomim)
dim(allphecode) #1534

ggplot(allphecode, aes(x=phecode, fill=source)) +geom_histogram(bins=40) + facet_wrap(~category, scales='free')

```

### Data descriptives 
``` {r, echo=F}
#number of genes
genedist=dataset %>% group_by(drugname) %>% distinct(entrez_id) %>% tally(name='gene.no')

plt1=ggplot(genedist, aes(x=gene.no))+ geom_histogram(binwidth = 2, color="black", fill="light blue") +
  labs(x='Number of gene targets',y='Number of drugs', title='Number of gene targets for drugs in dataset') +
  theme_classic()+
 geom_vline(aes(xintercept = mean(gene.no)),col='red',size=1) + annotate("text",color='red', x = 15, y = 300, label = paste('mean number of', '\n', 'gene targets =', round(mean(genedist$gene.no),2)))
  
#number of se
meanse=dataset %>% filter(!is.na(PHECODE_SE)) %>% group_by(drugname) %>% distinct(PHECODE_SE) %>% tally(name='se.no')

plt2=ggplot(meanse, aes(x=se.no))+ geom_histogram(binwidth = 2, color="black", fill="light blue") +
  labs(x='Number of phecode se',y='Number of drugs', title='Number of phecode se per drug in dataset') +
  theme_classic()+
 geom_vline(aes(xintercept = mean(se.no)),col='red',size=1) + annotate("text",color='red', x = 30, y = 100, label = paste('mean number of', '\n', 'phecode se =', round(mean(meanse$se.no),2)))

#mean tau
meantau=dataset %>% group_by(drugname) %>% distinct(Tau) %>% summarise(mean_tau=mean(Tau, na.rm=T))

plt3=ggplot(meantau, aes(x=mean_tau))+ geom_histogram(bins = 30, color="black", fill="light blue") +
  labs(x='Mean Tau score per drug',y='Number of drugs', title='Tau score per drug') +
  theme_classic()+
 geom_vline(aes(xintercept = mean(mean_tau)),col='red',size=1) + annotate("text",color='red', x = .5, y = 50, label = paste('mean tau =', round(mean(meantau$mean_tau),2)))

#mean constraint. #OE scores <1? 

meanoe=dataset %>% group_by(drugname) %>% distinct(oe_lof_upper) %>% summarise(mean_oe=mean(oe_lof_upper, na.rm=T))

plt4=ggplot(meanoe, aes(x=mean_oe))+ geom_histogram(bins = 30, color="black", fill="light blue") +
  labs(x='Mean oe score per drug',y='Number of drugs', title='oe score per drug') +
  theme_classic()+
 geom_vline(aes(xintercept = mean(mean_oe, na.rm = T)),col='red',size=1) + annotate("text",color='red', x = 1, y = 50, label = paste('mean oe =', round(mean(meanoe$mean_oe, na.rm = T),2)))

#number of omim phecodes  
meanomim=dataset %>% filter(!is.na(omim_phecode)) %>% group_by(drugname) %>% distinct(omim_phecode) %>% tally(name='omim.no')

plt5=ggplot(meanomim, aes(x=omim.no))+ geom_histogram(binwidth = 1, color="black", fill="light blue") +
  labs(x='Number of phecode omim phenotypes',y='Number of drugs', title='Number of phecode omim phenotypes per drug in dataset') +
  theme_classic()+
 geom_vline(aes(xintercept = mean(omim.no)),col='red',size=1) + annotate("text",color='red', x = 6, y = 270, label = paste('mean number of phecode', '\n', ' omim phenotypes=', round(mean(meanomim$omim.no),2)))

#Number of genes which had NA for omim
missingomim=dataset %>% distinct(drugname, entrez_id, omim_phecode) %>% filter(is.na(omim_phecode)) %>% group_by(drugname) %>% tally(name='no_omim_genes.no')

plt6=ggplot(missingomim, aes(x=no_omim_genes.no))+ geom_histogram(binwidth = 1, color="black", fill="light blue") +
  labs(x='Number of genes with NA for omim',y='Number of drugs', title='Number of genes which had NA for omim') +
  theme_classic()+
 geom_vline(aes(xintercept = mean(no_omim_genes.no)),col='red',size=1) + annotate("text",color='red', x = 13, y = 150, label = paste('mean number of','\n', 'genes =', round(mean(missingomim$no_omim_genes.no),2)))



#number of coloc phecodes 

meancolocpheno=dataset %>% filter(!is.na(coloc08_phecode)) %>% group_by(drugname) %>% distinct(coloc08_phecode) %>% tally(name='colocpheno.no')

plt7=ggplot(meancolocpheno, aes(x=colocpheno.no))+ geom_histogram(binwidth = 10, color="black", fill="light blue") +
  labs(x='Number of phecode coloc2 phenotypes',y='Number of drugs', title='Number of phecode coloc2  phenotypes per drug in dataset') +
  theme_classic()+
 geom_vline(aes(xintercept = mean(colocpheno.no)),col='red',size=1) + annotate("text",color='red', x = 170, y = 85, label = paste('mean number of phecode ', '\n', 'coloc2 phenotypes =', round(mean(meancolocpheno$colocpheno.no),2)))

meancolocphenotissue=dataset %>% group_by(coloc08_phecode) %>% distinct(Tissue) %>% tally(name='coloctissue.no')

plt8=ggplot(meancolocphenotissue, aes(x=coloctissue.no))+ geom_histogram(binwidth = 2, color="black", fill="light blue") +
  labs(x='Number of tissues that match to coloc2 phecode',y='Number of Tissues', title='Number of tissues mapped to coloc2 phenotypes') +
  theme_classic()+
 geom_vline(aes(xintercept = mean(coloctissue.no)),col='red',size=1) + annotate("text",color='red', x = 37,y  = 100, label = paste('mean number of tissues  ', '\n', 'per coloc phecode =', round(mean(meancolocphenotissue$coloctissue.no),2)))

#Number of genes per drug which had NA for coloc
missingcoloc=dataset %>% distinct(drugname, entrez_id, coloc08_phecode) %>% filter(is.na(coloc08_phecode)) %>% group_by(drugname) %>% tally(name='no_coloc_genes.no')

plt9=ggplot(missingcoloc, aes(x=no_coloc_genes.no))+ geom_histogram(binwidth = 1, color="black", fill="light blue") +
  labs(x='Number of genes with NA for coloc',y='Number of drugs', title='Number of genes per drug which had NA for coloc') +
  theme_classic()+
 geom_vline(aes(xintercept = mean(no_coloc_genes.no)),col='red',size=1) + annotate("text",color='red', x = 3, y = 120, label = paste('mean number of','\n', 'genes =', round(mean(missingcoloc$no_coloc_genes.no),2)))


grid.arrange(plt1,plt3,plt4)
grid.arrange(plt2,plt7,plt5)
grid.arrange(plt6,plt9,plt8)

descriptive_table=cbind.data.frame('No of drugs'=length(unique(dataset$drugname)),
                                   'Average no. genes'=mean(genedist$gene.no))


descriptive_table=rbind.data.frame('Drug gene targets'=cbind.data.frame(Unique_Number=paste(length(unique(dataset$entrez_id))), Mean=mean(genedist$gene.no), Std=sd(genedist$gene.no)),
                                   'Tau'=cbind.data.frame(Unique_Number=paste('-'), Mean=mean(meantau$mean_tau), Std=sd(meantau$mean_tau)),
                                   'Constraint'=cbind.data.frame(Unique_Number=paste('-'), Mean=mean(meanoe$mean_oe,na.rm  =T), Std=sd(meanoe$mean_oe,na.rm  =T)),
                                   'Side effect'=cbind.data.frame(Unique_Number=paste(length(unique(dataset$PHECODE_SE))), Mean=mean(meanse$se.no), Std=sd(meanse$se.no)),
                                   'Omim'=cbind.data.frame(Unique_Number=paste(length(unique(dataset$omim_phecode))), Mean=mean(meanomim$omim.no), Std=sd(meanomim$omim.no)),
                                   'Coloc Phenotype'=cbind.data.frame(Unique_Number=paste(length(unique(dataset$coloc08_phecode))), Mean=mean(meancolocpheno$colocpheno.no), Std=sd(meancolocpheno$colocpheno.no))
                                   )
descriptive_table

         #         Unique_Number       Mean        Std
# Drug gene targets           658  4.9796265  6.4772962
# Tau                           -  0.6803556  0.1818899
# Constraint                    -  0.7453552  0.3432504
# Side effect                 672 13.4567063 13.5091394
# Omim                         97  1.8666667  1.5925783
# Coloc Phenotype             765 80.6666667 66.3358518

```

### Dataset for model
1. Dataset - 589 unique drugs. 2933 unique drug-gene pairs X 672 unique SE = 1,970,976 lines

``` {r, echo=F}

se_phecodes=unique(dataset$PHECODE_SE)

dataset1=dataset[c(-9,-11)]
dim(dataset1)
# 7293455      10
drug_info=dataset[c(1:3)]
drug_info=drug_info[!duplicated(drug_info),]
dim(drug_info)
dataset_by_phecode=dataset1[c(1:3,4,8,9)]

#0 if gene doesnt match phecode se. NA if omim or coloc never reported gene
#Dataset- 589 unique drugs. 2933 unique drug-gene pairs X 672 unique SE
 
Byphecode2<-lapply(c(se_phecodes), function(se){
  #cat(paste(se,'\n'))
  dataset_phecode=dataset_by_phecode[which(dataset_by_phecode$PHECODE_SE==se),]
  drug_info$PheCode=se
  drug_info_one_all=merge(drug_info,dataset_phecode, by= paste(colnames(drug_info[-4])),all=T)
  drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 #cat(nrow(drug_info_one_all))
    drug_info_one_all[c(5:7)]<-lapply(drug_info_one_all[c(5:7)], function(col) ifelse(col==se,1,0))
   drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 drug_info_one_al2=drug_info_one_all[order(drug_info_one_all$PHECODE_SE,drug_info_one_all$coloc08_phecode,drug_info_one_all$omim_phecode, decreasing=T),]
 drug_info_one_al2$PHECODE_SE[is.na(drug_info_one_al2$PHECODE_SE)]<-0
 drug_info_one_al2=drug_info_one_al2[!duplicated(drug_info_one_al2[c(1:4)]),]
  # cat(paste(nrow(drug_info_one_al2)),'\n')
 drug_info_one_al2=drug_info_one_al2
})
Byphecode2<-do.call('rbind',Byphecode2)  
#paste(length(unique(Byphecode2$PheCode)))
dim(Byphecode2) #1970976
print('0 if gene doesnt match phecode se. NA if omim or coloc never reported gene')
lapply(Byphecode2[c(5:7)], function(x) summary(as.factor(x)))
#replaced all na for 0
Byphecode3=Byphecode2
Byphecode3[is.na(Byphecode3)]<-0

df=lapply(Byphecode3[c(5:7)], function(x) summary(as.factor(x)))
df=do.call('rbind',df)
df
chisq_se.coloc=chisq.test(df[-2,],correct = FALSE)
chisq_se.coloc
chisq_se.omim=chisq.test(df[-3,],correct = FALSE)
chisq_se.omim

# $PHECODE_SE
#       0       1 
# 1928048   42928 

# $omim_phecode
#       0       1 
# 1970951      25 

# $coloc08_phecode
#       0       1 
# 1970245     731 
print('Number of phecode se, coloc and omim matches per drug-gene pair')
 descriptivetable2druggene=Byphecode3%>% group_by(drugname,entrez_id) %>% summarise(total_se=sum(PHECODE_SE), total_coloc=sum(coloc08_phecode), total_omim=sum(omim_phecode))

head(descriptivetable2druggene)
descriptivetable2=lapply(c('se','coloc','omim'), function(x){
   column=paste0('total_',x)
   table=rbind.data.frame(x=cbind.data.frame(total=sum(descriptivetable2druggene[[column]]),
                                                                           percentage=round(sum((descriptivetable2druggene[[column]])/nrow(Byphecode3)*100),2),
                                                                           Mean=round(mean(descriptivetable2druggene[[column]]),2),
                                                                           Std=round(sd(descriptivetable2druggene[[column]]),2)))
   rownames(table)=paste(x, 'phecode trait per DG pair')
      table                    
 })

descriptivetable2<-do.call('rbind',descriptivetable2)
descriptivetable2
#                                total percentage  Mean   Std
# se phecode trait per DG pair    42928       2.18 14.64 14.63
# coloc phecode trait per DG pair   731       0.04  0.25  0.63
# omim phecode trait per DG pair     25       0.00  0.01  0.09
```
 1.logistic model
 2.mixed model
```{r}
logmod_coloc=glm(PHECODE_SE~coloc08_phecode, data = Byphecode3, family = 'binomial')
summary(logmod_coloc)
# Coefficients:
#                  Estimate Std. Error  z value Pr(>|z|)    
# (Intercept)     -3.821914   0.004921 -776.641   <2e-16 ***
# coloc08_phecode 17.387981  19.802898    0.878     0.38    

logmod_omim=glm(PHECODE_SE~omim_phecode, data = Byphecode3, family = 'binomial')
summary(logmod_omim)
#Coefficients:
#               Estimate Std. Error  z value Pr(>|z|)    
# (Intercept)  -3.805322   0.004881 -779.572   <2e-16 ***
# omim_phecode 14.371349  23.893609    0.601    0.548    

se_predictbycoloc=glmer(PHECODE_SE~coloc08_phecode+(1|PheCode), data = Byphecode3, family = 'binomial')
summary(se_predictbycoloc) #SIGNIFICANT
# Fixed effects:
#                 Estimate Std. Error z value Pr(>|z|)    
# (Intercept)     -4.85417    0.04635 -104.74   <2e-16 ***
# coloc08_phecode 24.64796    2.15747   11.42   <2e-16 ***
se_predictbyomim=glmer(PHECODE_SE~omim_phecode+(1|PheCode), data = Byphecode3, family = 'binomial')
summary(se_predictbyomim)
# Fixed effects:
#              Estimate Std. Error  z value Pr(>|z|)    
# (Intercept)  -4.83532    0.02612 -185.103   <2e-16 ***
# omim_phecode 20.12196   16.00034    1.258    0.209  
multi=glm(PHECODE_SE~coloc08_phecode+omim_phecode, data = Byphecode3, family = 'binomial')
summary(multi)


# Coefficients:
#                   Estimate Std. Error  z value Pr(>|z|)    
# (Intercept)      -3.822507   0.004923 -776.537   <2e-16 ***
# coloc08_phecode  17.388573  19.802898    0.878    0.380    
# omim_phecode     17.388573 107.082233    0.162    0.871 
```
2. Dataset - 589 unique drugs X 672 unique SE = 395,808 lines
All genes aggregated per drug
``` {r, echo=F}

##group all genes togther 

#Dataset- 589 unique drugs X 672 unique SE
#395,808 lines  

Byphecode_agggregategene<-lapply(c(se_phecodes), function(se){
 # cat(paste(se,'\n'))
  dataset_phecode=dataset_by_phecode[which(dataset_by_phecode$PHECODE_SE==se),]
  drug_info$PheCode=se
  drug_info_one_all=merge(drug_info,dataset_phecode, by= paste(colnames(drug_info[-4])),all=T)
  drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 #cat(nrow(drug_info_one_all))
    drug_info_one_all[c(5:7)]<-lapply(drug_info_one_all[c(5:7)], function(col) ifelse(col==se,1,0))
    drug_info_one_all[is.na(drug_info_one_all)]<-0
   drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
   drug_info_one_all3=aggregate(.~drugname, drug_info_one_all[c(-2,-3,-4)], sum)
   drug_info_one_all3$Phecode=se
   drug_info_one_all3[c(-1,-5)][drug_info_one_all3[c(-1,-5)]>1]<-1
  #  cat(paste(nrow(drug_info_one_all3)),'\n')
 drug_info_one_all3=drug_info_one_all3
})
Byphecode_agggregategene<-do.call('rbind',Byphecode_agggregategene)  
head(Byphecode_agggregategene)
df_1=lapply(Byphecode_agggregategene[c(2:4)], function(x) summary(as.factor(x)))
df_1=do.call('rbind',df_1)
df_1
chisq_se.coloc1=chisq.test(df_1[-2,],correct = FALSE)
chisq_se.coloc1
chisq_se.omim1=chisq.test(df_1[-3,],correct = FALSE)
chisq_se.omim1

print('Number of unique phecode se, coloc and omim matches per drug')

descriptivetable2druggene2=Byphecode_agggregategene%>% group_by(drugname) %>% summarise(total_se=sum(PHECODE_SE), total_coloc=sum(coloc08_phecode), total_omim=sum(omim_phecode))
head(descriptivetable2druggene2)
 
 descriptivetable_perdrug=lapply(c('se','coloc','omim'), function(x){
   column=paste0('total_',x)
   table=rbind.data.frame(x=cbind.data.frame(total=sum(descriptivetable2druggene2[[column]]),                                                                           percentage=round(sum((descriptivetable2druggene2[[column]])/nrow(Byphecode_agggregategene)*100),2),
                                                                           Mean=round(mean(descriptivetable2druggene2[[column]]),2),
                                                                           Std=round(sd(descriptivetable2druggene2[[column]]),2)))
   rownames(table)=paste(x, 'phecode trait per drug')
      table                    
 })

descriptivetable_perdrug<-do.call('rbind',descriptivetable_perdrug)
descriptivetable_perdrug


#                              total percentage  Mean   Std
# se phecode trait per drug     7926       2.00 13.46 13.51
# coloc phecode trait per drug   490       0.12  0.83  1.65
# omim phecode trait per drug     24       0.01  0.04  0.21
```
1.logistic model
2.mixed model
```{r}

logmod_coloc_drug=glm(PHECODE_SE~coloc08_phecode, data = Byphecode_agggregategene, family = 'binomial')
summary(logmod_coloc_drug)
#Coefficients:
#                 Estimate Std. Error  z value Pr(>|z|)    
# (Intercept)     -3.95437    0.01171 -337.772   <2e-16 ***
# coloc08_phecode 19.52044   65.74816    0.297    0.767    
logmod_omim_drug=glm(PHECODE_SE~omim_phecode, data = Byphecode_agggregategene, family = 'binomial')
summary(logmod_omim_drug)
#Coefficients:
#              Estimate Std. Error  z value Pr(>|z|)    
# (Intercept)  -3.89359    0.01136 -342.641   <2e-16 ***
# omim_phecode 16.45965   66.28803    0.248    0.804  
se_predictbycoloc_groupgene=glmer(PHECODE_SE~coloc08_phecode+(1|Phecode), family = 'binomial',data = Byphecode_agggregategene)
summary(se_predictbycoloc_groupgene)
#unable to evaluate scaled gradientModel failed to converge: degenerate  Hessian with 1 negative eigenvaluesvariance-covariance matrix computed from finite-difference #Hessian is
#not positive definite or contains NA values: falling back to var-cov estimated from RXvariance-covariance matrix computed from finite-difference Hessian is
#not positive definite or contains NA values: falling back to var-cov estimated from RXGeneralized linear mixed model fit by maximum likelihood (Laplace Approximation) 
# Fixed effects:
#                   Estimate Std. Error z value Pr(>|z|)    
# (Intercept)     -4.635e+00  5.129e-02 -90.363   <2e-16 ***
# coloc08_phecode  1.137e+04  3.032e+06   0.004    0.997    
se_predictbyomim_groupgene=glmer(PHECODE_SE~omim_phecode+(1|Phecode), family = 'binomial', data = Byphecode_agggregategene )
summary(se_predictbyomim_groupgene)
#Model is nearly unidentifiable: large eigenvalue ratio
 #- Rescale variables?Generalized linear mixed model fit by maximum likelihood (Laplace Approximation) ['glmerMod']

#Fixed effects:
 #            Estimate Std. Error z value Pr(>|z|)    
#(Intercept)  -4.56715    0.04853 -94.115   <2e-16 ***
#omim_phecode 26.85119   45.25484   0.593    0.553    
```

