---
title: "Drug_dataset_V4"
author: "Aine Duffy"
date: "7/01/2020"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


  ##Remove to broad phecodes? - Use only billable codes?
```{r, include=F}

library(data.table)
library(stringr)
library(xlsx)
library(dplyr)
library(lme4)

```


``` {r, echo=F}
dataset=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data/Dataset/drug_sider2_omim_tau_coloc_editedv4.txt',data.table=F)
#Phecode hierarchy 
phecode_dat1=unique(dataset[,4])

```


### All Phecodes
1. Dataset - 590 unique drugs. 2939 unique drug-gene pairs X 701 unique SE = 2,060,239 lines

``` {r, echo=F}

se_phecodes=unique(dataset$PHECODE_SE)

dataset1=dataset[c(-9,-11)]
dim(dataset1)
# 7258727      10
drug_info=dataset %>% distinct(drugname, entrez_id, symbol)
dim(drug_info)
dataset_by_phecode=dataset1[c(1:3,4,8,9)]
dataset_by_phecode=dataset_by_phecode[!duplicated(dataset_by_phecode),]
#0 if gene doesnt match phecode se. NA if omim or coloc never reported gene
#Dataset- 589 unique drugs. 2933 unique drug-gene pairs X 672 unique SE
 
Byphecode2<-lapply(c(se_phecodes), function(se){
  #cat(paste(se,'\n'))
  dataset_phecode=dataset_by_phecode[which(dataset_by_phecode$PHECODE_SE==se),]
  drug_info$PheCode=se
  drug_info_one_all=merge(drug_info,dataset_phecode, by= paste(colnames(drug_info[-4])),all=T)
  drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 #cat(nrow(drug_info_one_all))
    drug_info_one_all[c(5:7)]<-lapply(drug_info_one_all[c(5:7)], function(col) ifelse(col==se,1,0))
   drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 drug_info_one_al2=drug_info_one_all[order(drug_info_one_all$PHECODE_SE,drug_info_one_all$coloc08_phecode,drug_info_one_all$omim_phecode, decreasing=T),]
 drug_info_one_al2$PHECODE_SE[is.na(drug_info_one_al2$PHECODE_SE)]<-0
 drug_info_one_al2=drug_info_one_al2[!duplicated(drug_info_one_al2[c(1:4)]),]
  # cat(paste(nrow(drug_info_one_al2)),'\n')
 drug_info_one_al2=drug_info_one_al2
})
Byphecode2<-do.call('rbind',Byphecode2)  

Byphecode3=Byphecode2
Byphecode3[is.na(Byphecode3)]<-0
head(Byphecode3)
lapply(Byphecode3[c(5:7)], function(x) summary(as.factor(x)))
```

Number of phecode se, coloc and omim matches per drug-gene pair
```{r,echo=F}
 descriptivetable2druggene=Byphecode3%>% group_by(drugname,entrez_id) %>% summarise(total_se=sum(PHECODE_SE), total_coloc=sum(coloc08_phecode), total_omim=sum(omim_phecode))

head(descriptivetable2druggene)
descriptivetable2=lapply(c('se','coloc','omim'), function(x){
   column=paste0('total_',x)
   table=rbind.data.frame(x=cbind.data.frame(total=sum(descriptivetable2druggene[[column]]),
                                                                           percentage=round(sum((descriptivetable2druggene[[column]])/nrow(Byphecode3)*100),2),
                                                                           Mean=round(mean(descriptivetable2druggene[[column]]),2),
                                                                           Std=round(sd(descriptivetable2druggene[[column]]),2)))
   rownames(table)=paste(x, 'phecode trait per DG pair')
      table                    
 })

descriptivetable2<-do.call('rbind',descriptivetable2)
descriptivetable2
```

#### Chisq test
1.Phecode se and OMIM phecode 
```{r, echo=F}
################################# chisq test all dg pairs ##################################################################
#omim
Datasetv3_dg_omim=rbind.data.frame('Phecode se yes'=(cbind('omim phecode Yes'=length(Byphecode3$PHECODE_SE[Byphecode3$PHECODE_SE==1 & Byphecode3$omim_phecode==1]),'omim phecode No'= length(Byphecode3$PHECODE_SE[Byphecode3$PHECODE_SE==1 & Byphecode3$omim_phecode==0]))),
'Phecode se no'=(cbind('omim phecode Yes'=length(Byphecode3$PHECODE_SE[Byphecode3$PHECODE_SE==0 & Byphecode3$omim_phecode==1]),'omim phecode No'= length(Byphecode3$PHECODE_SE[Byphecode3$PHECODE_SE==0 & Byphecode3$omim_phecode==0]))))
Datasetv3_dg_omim
D3_chisq_se.omim=chisq.test(Datasetv3_dg_omim,correct = FALSE)
D3_chisq_se.omim
#'Warning message: In chisq.test(Datasetv3_dg_omim, correct = FALSE)   Chi-squared approximation may be incorrect'
D3_chisq_se.omim=chisq.test(Datasetv3_dg_omim,correct = FALSE,simulate.p.value = TRUE)
D3_chisq_se.omim
```

2.Phecode se and coloc phecode 
```{r,echo=F}
#coloc
Datasetv3_dg_coloc=rbind.data.frame('Phecode se yes'=(cbind('coloc phecode Yes'=length(Byphecode3$PHECODE_SE[Byphecode3$PHECODE_SE==1 & Byphecode3$coloc08_phecode==1]),'coloc phecode No'= length(Byphecode3$PHECODE_SE[Byphecode3$PHECODE_SE==1 & Byphecode3$coloc08_phecode==0]))),
                                    'Phecode se no'=(cbind('coloc phecode Yes'=length(Byphecode3$PHECODE_SE[Byphecode3$PHECODE_SE==0 & Byphecode3$coloc08_phecode==1]),'coloc phecode No'= length(Byphecode3$PHECODE_SE[Byphecode3$PHECODE_SE==0 & Byphecode3$coloc08_phecode==0]))))

Datasetv3_dg_coloc
D3_chisq_se.coloc=chisq.test(Datasetv3_dg_coloc,correct = FALSE)
D3_chisq_se.coloc
D3_chisq_se.coloc=chisq.test(Datasetv3_dg_coloc,correct = FALSE,simulate.p.value = TRUE)
D3_chisq_se.coloc



```
 
#### Models 
 1.logistic model
```{r}
logmod_coloc=glm(PHECODE_SE~coloc08_phecode, data = Byphecode3, family = 'binomial')
summary(logmod_coloc)

logmod_omim=glm(PHECODE_SE~omim_phecode, data = Byphecode3, family = 'binomial')
summary(logmod_omim)
 
```
 
 2.mixed model
```{r,echo=F}
se_predictbycoloc=glmer(PHECODE_SE~coloc08_phecode+(1|PheCode), data = Byphecode3, family = 'binomial')
summary(se_predictbycoloc) 

se_predictbyomim=glmer(PHECODE_SE~omim_phecode+(1|PheCode), data = Byphecode3, family = 'binomial')
summary(se_predictbyomim)

multi=glm(PHECODE_SE~coloc08_phecode+omim_phecode, data = Byphecode3, family = 'binomial')
summary(multi)

```

### All Phecodes. Genes collapsed
2. Dataset - 590 unique drugs X 701 unique SE = 395,808 lines
All genes aggregated per drug
``` {r, echo=F}
##group all genes togther 

#Dataset- 589 unique drugs X 672 unique SE
#395,808 lines  

Byphecode_agggregategene<-lapply(c(se_phecodes), function(se){
 # cat(paste(se,'\n'))
  dataset_phecode=dataset_by_phecode[which(dataset_by_phecode$PHECODE_SE==se),]
  drug_info$PheCode=se
  drug_info_one_all=merge(drug_info,dataset_phecode, by= paste(colnames(drug_info[-4])),all=T)
  drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 #cat(nrow(drug_info_one_all))
    drug_info_one_all[c(5:7)]<-lapply(drug_info_one_all[c(5:7)], function(col) ifelse(col==se,1,0))
    drug_info_one_all[is.na(drug_info_one_all)]<-0
   drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
   drug_info_one_all3=aggregate(.~drugname, drug_info_one_all[c(-2,-3,-4)], sum)
   drug_info_one_all3$PheCode=se
   drug_info_one_all3[c(-1,-5)][drug_info_one_all3[c(-1,-5)]>1]<-1
  #  cat(paste(nrow(drug_info_one_all3)),'\n')
 drug_info_one_all3=drug_info_one_all3
})
Byphecode_agggregategene<-do.call('rbind',Byphecode_agggregategene)  
head(Byphecode_agggregategene)
lapply(Byphecode_agggregategene[c(2:4)], function(x) summary(as.factor(x)))

```
Number of unique phecode se, coloc and omim matches per drug
```{r,echo=F}

descriptivetable2druggene2=Byphecode_agggregategene%>% group_by(drugname) %>% summarise(total_se=sum(PHECODE_SE), total_coloc=sum(coloc08_phecode), total_omim=sum(omim_phecode))
head(descriptivetable2druggene2)
 
 descriptivetable_perdrug=lapply(c('se','coloc','omim'), function(x){
   column=paste0('total_',x)
   table=rbind.data.frame(x=cbind.data.frame(total=sum(descriptivetable2druggene2[[column]]),                                                                           percentage=round(sum((descriptivetable2druggene2[[column]])/nrow(Byphecode_agggregategene)*100),2),
                                                                           Mean=round(mean(descriptivetable2druggene2[[column]]),2),
                                                                           Std=round(sd(descriptivetable2druggene2[[column]]),2)))
   rownames(table)=paste(x, 'phecode trait per drug')
      table                    
 })

descriptivetable_perdrug<-do.call('rbind',descriptivetable_perdrug)
descriptivetable_perdrug

``` 

#### Chisq test
1.Phecode se and OMIM phecode
``` {r,echo=F}
################################# chisq test aggregate dg pairs ##################################################################

#omim
Datasetv3_dg_omim_aggregate=rbind.data.frame('Phecode se yes'=(cbind('omim phecode Yes'=length(Byphecode_agggregategene$PHECODE_SE[Byphecode_agggregategene$PHECODE_SE==1 & Byphecode_agggregategene$omim_phecode==1]),'omim phecode No'= length(Byphecode_agggregategene$PHECODE_SE[Byphecode_agggregategene$PHECODE_SE==1 & Byphecode_agggregategene$omim_phecode==0]))),
                                             'Phecode se no'=(cbind('omim phecode Yes'=length(Byphecode_agggregategene$PHECODE_SE[Byphecode_agggregategene$PHECODE_SE==0 & Byphecode_agggregategene$omim_phecode==1]),'omim phecode No'= length(Byphecode_agggregategene$PHECODE_SE[Byphecode_agggregategene$PHECODE_SE==0 & Byphecode_agggregategene$omim_phecode==0]))))

D3ag_chisq_se.omim=chisq.test(Datasetv3_dg_omim_aggregate,correct = FALSE)
D3ag_chisq_se.omim
D3ag_chisq_se.omim=chisq.test(Datasetv3_dg_omim_aggregate,correct = FALSE,simulate.p.value = TRUE)
D3ag_chisq_se.omim
```

2.Phecode se and coloc phecode 
```{r,echo=F}
#coloc

Datasetv3_dg_coloc_aggregate=rbind.data.frame('Phecode se yes'=(cbind('coloc phecode Yes'=length(Byphecode_agggregategene$PHECODE_SE[Byphecode_agggregategene$PHECODE_SE==1 & Byphecode_agggregategene$coloc08_phecode==1]),'coloc phecode No'= length(Byphecode_agggregategene$PHECODE_SE[Byphecode_agggregategene$PHECODE_SE==1 & Byphecode_agggregategene$coloc08_phecode==0]))),
                                              'Phecode se no'=(cbind('coloc phecode Yes'=length(Byphecode_agggregategene$PHECODE_SE[Byphecode_agggregategene$PHECODE_SE==0 & Byphecode_agggregategene$coloc08_phecode==1]),'coloc phecode No'= length(Byphecode_agggregategene$PHECODE_SE[Byphecode_agggregategene$PHECODE_SE==0 & Byphecode_agggregategene$coloc08_phecode==0]))))

D3ag_chisq_se.coloc=chisq.test(Datasetv3_dg_coloc_aggregate,correct = FALSE)
D3ag_chisq_se.coloc
D3ag_chisq_se.coloc=chisq.test(Datasetv3_dg_coloc_aggregate,correct = FALSE,simulate.p.value = TRUE)
D3ag_chisq_se.coloc
```
#### Models
1.logistic model
```{r}

logmod_coloc_drug=glm(PHECODE_SE~coloc08_phecode, data = Byphecode_agggregategene, family = 'binomial')
summary(logmod_coloc_drug)
  
logmod_omim_drug=glm(PHECODE_SE~omim_phecode, data = Byphecode_agggregategene, family = 'binomial')
summary(logmod_omim_drug)
```

2.Phecode se and coloc phecode 
```{r,echo=F}
se_predictbycoloc_groupgene=glmer(PHECODE_SE~coloc08_phecode+(1|PheCode), family = 'binomial',data = Byphecode_agggregategene)
summary(se_predictbycoloc_groupgene)
 
se_predictbyomim_groupgene=glmer(PHECODE_SE~omim_phecode+(1|PheCode), family = 'binomial', data = Byphecode_agggregategene )
summary(se_predictbyomim_groupgene)
 
```

### Grouped Phecode child terms by lowest level
1. Dataset - 590 unique drugs. 2939 unique drug-gene pairs X 651 unique SE = 2,060,239 lines

``` {r, echo=F}
dataset2=dataset[c(-9,-11)]
dataset2$PHECODE_SE_grouped1=ifelse(!is.na(str_extract(dataset2$PHECODE_SE,'\\.[0-9]')), paste0(0,str_extract(dataset2$PHECODE_SE,'\\.[0-9]')), 0)
dataset2$PHECODE_SE_grouped=trunc(dataset2$PHECODE_SE)+as.numeric(dataset2$PHECODE_SE_grouped1)
head(dataset2)
dataset2$coloc08_phecode_grouped1=ifelse(!is.na(str_extract(dataset2$coloc08_phecode,'\\.[0-9]')), paste0(0,str_extract(dataset2$coloc08_phecode,'\\.[0-9]')), 0)
dataset2$coloc08_phecode_grouped=trunc(dataset2$coloc08_phecode)+as.numeric(dataset2$coloc08_phecode_grouped1)

dataset2$omim_phecode_grouped1=ifelse(!is.na(str_extract(dataset2$omim_phecode,'\\.[0-9]')), paste0(0,str_extract(dataset2$omim_phecode,'\\.[0-9]')), 0)
dataset2$omim_phecode_grouped=trunc(dataset2$omim_phecode)+as.numeric(dataset2$omim_phecode_grouped1)

dim(dataset2)
# 7258727      10
drug_info=dataset %>% distinct(drugname, entrez_id, symbol)
dim(drug_info)
dataset_by_phecode_collapsed=dataset2[c(1:3,12,14,16)]
dataset_by_phecode_collapsed=dataset_by_phecode_collapsed[!duplicated(dataset_by_phecode_collapsed),]
#0 if gene doesnt match phecode se. NA if omim or coloc never reported gene
#Dataset- 589 unique drugs. 2933 unique drug-gene pairs X 672 unique SE
se_phecodes_collapsed=unique(dataset_by_phecode_collapsed$PHECODE_SE_grouped)
 
Byphecode2_collapsed<-lapply(c(se_phecodes_collapsed), function(se){
  #cat(paste(se,'\n'))
  dataset_phecode=dataset_by_phecode_collapsed[which(dataset_by_phecode_collapsed$PHECODE_SE_grouped==se),]
  drug_info$PheCode=se
  drug_info_one_all=merge(drug_info,dataset_phecode, by= paste(colnames(drug_info[-4])),all=T)
  drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 #cat(nrow(drug_info_one_all))
    drug_info_one_all[c(5:7)]<-lapply(drug_info_one_all[c(5:7)], function(col) ifelse(col==se,1,0))
   drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 drug_info_one_al2=drug_info_one_all[order(drug_info_one_all$PHECODE_SE_grouped,drug_info_one_all$coloc08_phecode_grouped,drug_info_one_all$omim_phecode_grouped, decreasing=T),]
 drug_info_one_al2$PHECODE_SE_grouped[is.na(drug_info_one_al2$PHECODE_SE_grouped)]<-0
 drug_info_one_al2=drug_info_one_al2[!duplicated(drug_info_one_al2[c(1:4)]),]
  # cat(paste(nrow(drug_info_one_al2)),'\n')
 drug_info_one_al2=drug_info_one_al2
})
Byphecode2_collapsed<-do.call('rbind',Byphecode2_collapsed)  
#paste(length(unique(Byphecode2$PheCode)))
#dim(Byphecode2_collapsed) #1970976
#print('0 if gene doesnt match phecode se. NA if omim or coloc never reported gene')
#lapply(Byphecode2_collapsed[c(5:7)], function(x) summary(as.factor(x)))
#replaced all na for 0
Byphecode2_collapsed[is.na(Byphecode2_collapsed)]<-0
head(Byphecode2_collapsed)
lapply(Byphecode2_collapsed[c(5:7)], function(x) summary(as.factor(x)))

```

Number of phecode se, coloc and omim matches per drug-gene pair once child terms collapsed
```{r,echo=F}

 descriptivetable2druggene_collapsed=Byphecode2_collapsed%>% group_by(drugname,entrez_id) %>% summarise(total_se=sum(PHECODE_SE_grouped), total_coloc=sum(coloc08_phecode_grouped), total_omim=sum(omim_phecode_grouped))

head(descriptivetable2druggene_collapsed)
descriptivetable2_collapsed=lapply(c('se','coloc','omim'), function(x){
   column=paste0('total_',x)
   table=rbind.data.frame(x=cbind.data.frame(total=sum(descriptivetable2druggene_collapsed[[column]]),
                                                                           percentage=round(sum((descriptivetable2druggene_collapsed[[column]])/nrow(Byphecode2_collapsed)*100),2),
                                                                           Mean=round(mean(descriptivetable2druggene_collapsed[[column]]),2),
                                                                           Std=round(sd(descriptivetable2druggene_collapsed[[column]]),2)))
   rownames(table)=paste(x, 'phecode trait per DG pair')
      table                    
 })

descriptivetable2_collapsed<-do.call('rbind',descriptivetable2_collapsed)
descriptivetable2_collapsed

```

#### Chisq test
1.Phecode se and OMIM phecode
```{r,echo=F}
################################# chisq test all dg pairs ##################################################################
#omim
Dataset_dg_omim_collapse=rbind.data.frame('Phecode se yes'=(cbind('omim phecode Yes'=length(Byphecode2_collapsed$PHECODE_SE_grouped[Byphecode2_collapsed$PHECODE_SE_grouped==1 & Byphecode2_collapsed$omim_phecode_grouped==1]),'omim phecode No'= length(Byphecode2_collapsed$PHECODE_SE_grouped[Byphecode2_collapsed$PHECODE_SE_grouped==1 & Byphecode2_collapsed$omim_phecode_grouped==0]))),
'Phecode se no'=(cbind('omim phecode Yes'=length(Byphecode2_collapsed$PHECODE_SE_grouped[Byphecode2_collapsed$PHECODE_SE_grouped==0 & Byphecode2_collapsed$omim_phecode_grouped==1]),'omim phecode No'= length(Byphecode2_collapsed$PHECODE_SE_grouped[Byphecode2_collapsed$PHECODE_SE_grouped==0 & Byphecode2_collapsed$omim_phecode_grouped==0]))))
Dataset_dg_omim_collapse
D3_chisq_se.omim_collapse=chisq.test(Dataset_dg_omim_collapse,correct = FALSE)
D3_chisq_se.omim_collapse
#'Warning message: In chisq.test(Datasetv3_dg_omim, correct = FALSE)   Chi-squared approximation may be incorrect'
D3_chisq_se.omim_collapse=chisq.test(Dataset_dg_omim_collapse,correct = FALSE,simulate.p.value = TRUE)
D3_chisq_se.omim_collapse
```
1.Phecode se and coloc phecode
```{r,echo=F}
#coloc
Dataset_dg_coloc_collapse=rbind.data.frame('Phecode se yes'=(cbind('coloc phecode Yes'=length(Byphecode2_collapsed$PHECODE_SE_grouped[Byphecode2_collapsed$PHECODE_SE_grouped==1 & Byphecode2_collapsed$coloc08_phecode_grouped==1]),'coloc phecode No'= length(Byphecode2_collapsed$PHECODE_SE_grouped[Byphecode2_collapsed$PHECODE_SE_grouped==1 & Byphecode2_collapsed$coloc08_phecode_grouped==0]))),
                                    'Phecode se no'=(cbind('coloc phecode Yes'=length(Byphecode2_collapsed$PHECODE_SE_grouped[Byphecode2_collapsed$PHECODE_SE_grouped==0 & Byphecode2_collapsed$coloc08_phecode_grouped==1]),'coloc phecode No'= length(Byphecode2_collapsed$PHECODE_SE_grouped[Byphecode2_collapsed$PHECODE_SE_grouped==0 & Byphecode2_collapsed$coloc08_phecode_grouped==0]))))

Dataset_dg_coloc_collapse
D3_chisq_se.coloc_collapse=chisq.test(Dataset_dg_coloc_collapse,correct = FALSE)
D3_chisq_se.coloc_collapse
```
#### Models
1.logistic model 

```{r}
logmod_coloc_collapsed=glm(PHECODE_SE_grouped~coloc08_phecode_grouped, data = Byphecode2_collapsed, family = 'binomial')
summary(logmod_coloc_collapsed)

logmod_omim_collapsed=glm(PHECODE_SE_grouped~omim_phecode_grouped, data = Byphecode2_collapsed, family = 'binomial')
summary(logmod_omim_collapsed)
```

2.mixed model
```{r,echo=F}
se_predictbycoloc_collapsed=glmer(PHECODE_SE_grouped~coloc08_phecode_grouped+(1|PheCode), data = Byphecode2_collapsed, family = 'binomial')
summary(se_predictbycoloc_collapsed) 

se_predictbyomim_collapsed=glmer(PHECODE_SE_grouped~omim_phecode_grouped+(1|PheCode), data = Byphecode2_collapsed, family = 'binomial')
summary(se_predictbyomim_collapsed)

multi_collapsed=glm(PHECODE_SE_grouped~coloc08_phecode_grouped+omim_phecode_grouped, data = Byphecode2_collapsed, family = 'binomial')
summary(multi_collapsed)
```
### Grouped Phecode child terms by lowest level. Genes collapsed
2. Dataset - 590 unique drugs X 651 unique SE = 384,090 lines
All genes aggregated per drug
``` {r, echo=F}
##group all genes togther 

#Dataset- 589 unique drugs X 672 unique SE
#395,808 lines  

Byphecode_agggregategene_collapsed<-lapply(c(se_phecodes_collapsed), function(se){
 # cat(paste(se,'\n'))
  dataset_phecode=dataset_by_phecode_collapsed[which(dataset_by_phecode_collapsed$PHECODE_SE_grouped==se),]
  drug_info$PheCode=se
  drug_info_one_all=merge(drug_info,dataset_phecode, by= paste(colnames(drug_info[-4])),all=T)
  drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 #cat(nrow(drug_info_one_all))
    drug_info_one_all[c(5:7)]<-lapply(drug_info_one_all[c(5:7)], function(col) ifelse(col==se,1,0))
    drug_info_one_all[is.na(drug_info_one_all)]<-0
   drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
   drug_info_one_all3=aggregate(.~drugname, drug_info_one_all[c(-2,-3,-4)], sum)
   drug_info_one_all3$PheCode=se
   drug_info_one_all3[c(-1,-5)][drug_info_one_all3[c(-1,-5)]>1]<-1
  #  cat(paste(nrow(drug_info_one_all3)),'\n')
 drug_info_one_all3=drug_info_one_all3
})
Byphecode_agggregategene_collapsed<-do.call('rbind',Byphecode_agggregategene_collapsed)  
head(Byphecode_agggregategene_collapsed)
lapply(Byphecode_agggregategene_collapsed[c(2:4)], function(x) summary(as.factor(x)))

```
Number of unique phecode se, coloc and omim matches per drug once child terms collapsed
```{r,echo=F}

descriptivetable2druggene2_collapse=Byphecode_agggregategene_collapsed%>% group_by(drugname) %>% summarise(total_se=sum(PHECODE_SE_grouped), total_coloc=sum(coloc08_phecode_grouped), total_omim=sum(omim_phecode_grouped))
head(descriptivetable2druggene2_collapse)
 
 descriptivetable_perdrug_collapse=lapply(c('se','coloc','omim'), function(x){
   column=paste0('total_',x)
   table=rbind.data.frame(x=cbind.data.frame(total=sum(descriptivetable2druggene2_collapse[[column]]),                                                                           percentage=round(sum((descriptivetable2druggene2_collapse[[column]])/nrow(Byphecode_agggregategene_collapsed)*100),2),
                                                                           Mean=round(mean(descriptivetable2druggene2_collapse[[column]]),2),
                                                                           Std=round(sd(descriptivetable2druggene2_collapse[[column]]),2)))
   rownames(table)=paste(x, 'phecode trait per drug')
      table                    
 })

descriptivetable_perdrug_collapse<-do.call('rbind',descriptivetable_perdrug_collapse)
descriptivetable_perdrug_collapse
```

#### Chisq test
1.Phecode se and OMIM phecode
```{r,echo = F}
################################# chisq test aggregate dg pairs ##################################################################

#omim
Datasetv3_dg_omim_aggregate_collapse=rbind.data.frame('Phecode se yes'=(cbind('omim phecode Yes'=length(Byphecode_agggregategene_collapsed$PHECODE_SE_grouped[Byphecode_agggregategene_collapsed$PHECODE_SE_grouped==1 & Byphecode_agggregategene_collapsed$omim_phecode_grouped==1]),'omim phecode No'= length(Byphecode_agggregategene_collapsed$PHECODE_SE_grouped[Byphecode_agggregategene_collapsed$PHECODE_SE_grouped==1 & Byphecode_agggregategene_collapsed$omim_phecode_grouped==0]))),
                                             'Phecode se no'=(cbind('omim phecode Yes'=length(Byphecode_agggregategene_collapsed$PHECODE_SE_grouped[Byphecode_agggregategene_collapsed$PHECODE_SE_grouped==0 & Byphecode_agggregategene_collapsed$omim_phecode_grouped==1]),'omim phecode No'= length(Byphecode_agggregategene_collapsed$PHECODE_SE_grouped[Byphecode_agggregategene_collapsed$PHECODE_SE_grouped==0 & Byphecode_agggregategene_collapsed$omim_phecode_grouped==0]))))

D3ag_chisq_se.omim_collapse=chisq.test(Datasetv3_dg_omim_aggregate_collapse,correct = FALSE)
D3ag_chisq_se.omim_collapse
D3ag_chisq_se.omim_collapse=chisq.test(Datasetv3_dg_omim_aggregate_collapse,correct = FALSE,simulate.p.value = TRUE)
D3ag_chisq_se.omim_collapse
```

2.Phecode se and coloc phecode
```{r,echo=F}
#coloc

Datasetv3_dg_coloc_aggregate_collapse=rbind.data.frame('Phecode se yes'=(cbind('coloc phecode Yes'=length(Byphecode_agggregategene_collapsed$PHECODE_SE_grouped[Byphecode_agggregategene_collapsed$PHECODE_SE_grouped==1 & Byphecode_agggregategene_collapsed$coloc08_phecode_grouped==1]),'coloc phecode No'= length(Byphecode_agggregategene_collapsed$PHECODE_SE_grouped[Byphecode_agggregategene_collapsed$PHECODE_SE_grouped==1 & Byphecode_agggregategene_collapsed$coloc08_phecode_grouped==0]))),
                                              'Phecode se no'=(cbind('coloc phecode Yes'=length(Byphecode_agggregategene_collapsed$PHECODE_SE_grouped[Byphecode_agggregategene_collapsed$PHECODE_SE_grouped==0 & Byphecode_agggregategene_collapsed$coloc08_phecode_grouped==1]),'coloc phecode No'= length(Byphecode_agggregategene_collapsed$PHECODE_SE_grouped[Byphecode_agggregategene_collapsed$PHECODE_SE_grouped==0 & Byphecode_agggregategene_collapsed$coloc08_phecode_grouped==0]))))

D3ag_chisq_se.coloc_collapse=chisq.test(Datasetv3_dg_coloc_aggregate_collapse,correct = FALSE)
D3ag_chisq_se.coloc_collapse
D3ag_chisq_se.coloc_collapse=chisq.test(Datasetv3_dg_coloc_aggregate_collapse,correct = FALSE,simulate.p.value = TRUE)
D3ag_chisq_se.coloc_collapse
```

#### Models
1.logistic model

```{r}

logmod_coloc_drug_collapse=glm(PHECODE_SE_grouped~coloc08_phecode_grouped, data = Byphecode_agggregategene_collapsed, family = 'binomial')
summary(logmod_coloc_drug_collapse)
  
logmod_omim_drug_collapse=glm(PHECODE_SE_grouped~omim_phecode_grouped, data = Byphecode_agggregategene_collapsed, family = 'binomial')
summary(logmod_omim_drug_collapse)

```

2.mixed model

```{r,echo=F}
se_predictbycoloc_groupgene_collapse=glmer(PHECODE_SE_grouped~coloc08_phecode_grouped+(1|PheCode), family = 'binomial',data = Byphecode_agggregategene_collapsed)
summary(se_predictbycoloc_groupgene_collapse)
 
se_predictbyomim_groupgene_collapse=glmer(PHECODE_SE_grouped~omim_phecode_grouped+(1|PheCode), family = 'binomial', data = Byphecode_agggregategene_collapsed )
summary(se_predictbyomim_groupgene_collapse)

```

### Stratified
ALL
```{r,echo=F}
head(dataset2)
phecodef=read.csv('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/phecode_definitions1.2.csv')
#Byphecode_agggregategene_collapsed$category=phecodef$category[match(Byphecode_agggregategene_collapsed$Phecode, phecodef$phecode)]
Byphecode3$category=phecodef$category[match(Byphecode3$PheCode, phecodef$phecode)]
glmer(PHECODE_SE~coloc08_phecode+(1|PheCode), family = 'binomial',data = Byphecode3)

subseting<-unique(as.character(Byphecode3$category))

results_strat<-lapply(subseting, function(x){
  df=Byphecode3[Byphecode3$category==x,]
a<-glmer(PHECODE_SE~coloc08_phecode+(1|PheCode), data=df, family=binomial)
model1<-t(cbind.data.frame(
  summary(a)$coefficient[-1, c(1, 2, 4)]))
model<-(cbind(model1, warningmessage=paste(summary(a)$optinfo$conv$lme4$messages[1],summary(a)$optinfo$conv$lme4$messages[2])))
print(model)  
if(nrow(model)==0|ncol(model)==0){
   MODEL1<-as.data.frame(t(rep(NA,4)))
    rownames(MODEL1)<-x 
    colnames(MODEL1)<-c('Estimate', 'Std. Error',  'Pr(>|z|)','warningmessage')
}
  else if (nrow(model)!=0&ncol(model)==4) {
 print(paste(ncol(model), nrow(model)))
     rownames(model)<-x
  colnames(model)<-c('Estimate', 'Std. Error',  'Pr(>|z|)','warningmessage')
  MODEL1<-model}
else if (nrow(model)!=0&ncol(model)==3) {
 print(paste(ncol(model), nrow(model)))
  rownames(model)<-x
  model<-cbind(model,NA)
  colnames(model)<-c('Estimate', 'Std. Error',  'Pr(>|z|)','warningmessage')
  MODEL1<-model 
}
MODEL1
  
})

results_strat<-do.call('rbind',results_strat)
results_strat


```

Phecodes Collapsed

```{r,echo=F}

Byphecode2_collapsed$category=phecodef$category[match(Byphecode2_collapsed$PheCode, phecodef$phecode)]

subseting<-unique(as.character(Byphecode2_collapsed$category))

results_strat_phecodescollapsed<-lapply(subseting, function(x){
  df=Byphecode2_collapsed[Byphecode2_collapsed$category==x,]
a<-glmer(PHECODE_SE_grouped~coloc08_phecode_grouped+(1|PheCode), data=df, family=binomial)
model1<-t(cbind.data.frame(
  summary(a)$coefficient[-1, c(1, 2, 4)]))
model<-(cbind(model1, warningmessage=paste(summary(a)$optinfo$conv$lme4$messages[1],summary(a)$optinfo$conv$lme4$messages[2])))
print(model)  
if(nrow(model)==0|ncol(model)==0){
   MODEL1<-as.data.frame(t(rep(NA,4)))
    rownames(MODEL1)<-x 
    colnames(MODEL1)<-c('Estimate', 'Std. Error',  'Pr(>|z|)','warningmessage')
}
  else if (nrow(model)!=0&ncol(model)==4) {
 print(paste(ncol(model), nrow(model)))
     rownames(model)<-x
  colnames(model)<-c('Estimate', 'Std. Error',  'Pr(>|z|)','warningmessage')
  MODEL1<-model}
else if (nrow(model)!=0&ncol(model)==3) {
 print(paste(ncol(model), nrow(model)))
  rownames(model)<-x
  model<-cbind(model,NA)
  colnames(model)<-c('Estimate', 'Std. Error',  'Pr(>|z|)','warningmessage')
  MODEL1<-model 
}
MODEL1
  
})

results_strat_phecodescollapsed<-do.call('rbind',results_strat_phecodescollapsed)
results_strat_phecodescollapsed
```

### Main Indication. All Phecodes 
1. Dataset - 503 unique drugs. 2939 unique drug-gene pairs X 701 unique SE = 2,060,239 lines

```{r,echo=F}

mi_dataset=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data/Dataset/drug_sider2_omim_tau_coloc_MI.txt',data.table=F)


mi_phecodes=unique(mi_dataset$PHECODE_MI)

mi_dataset1=mi_dataset[c(-9,-11)]
dim(mi_dataset1)
# 7258727      10
drug_info_mi=mi_dataset1 %>% distinct(drugname, entrez_id, symbol)
dim(drug_info_mi)
dataset_by_miphecode=mi_dataset1[c(1:3,4,8,9)]
dataset_by_miphecode=dataset_by_miphecode[!duplicated(dataset_by_miphecode),]
#0 if gene doesnt match phecode se. NA if omim or coloc never reported gene
#Dataset- 589 unique drugs. 2933 unique drug-gene pairs X 672 unique SE
head(dataset_by_miphecode) 
Byphecode_mi2<-lapply(c(mi_phecodes), function(mi){
  #cat(paste(mi,'\n'))
  dataset_phecode=dataset_by_miphecode[which(dataset_by_miphecode$PHECODE_MI==mi),]
  drug_info_mi$PheCode=mi
  drug_info_one_all=merge(drug_info_mi,dataset_phecode, by= paste(colnames(drug_info_mi[-4])),all=T)
  drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 #cat(nrow(drug_info_one_all))
    drug_info_one_all[c(5:7)]<-lapply(drug_info_one_all[c(5:7)], function(col) ifelse(col==mi,1,0))
   drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 drug_info_one_al2=drug_info_one_all[order(drug_info_one_all$PHECODE_MI,drug_info_one_all$coloc08_phecode,drug_info_one_all$omim_phecode, decreasing=T),]
 drug_info_one_al2$PHECODE_MI[is.na(drug_info_one_al2$PHECODE_MI)]<-0
 drug_info_one_al2=drug_info_one_al2[!duplicated(drug_info_one_al2[c(1:4)]),]
  # cat(paste(nrow(drug_info_one_al2)),'\n')
 drug_info_one_al2=drug_info_one_al2
})
Byphecode_mi2<-do.call('rbind',Byphecode_mi2)  

#replaced all na for 0
Byphecode_mi3=Byphecode_mi2
Byphecode_mi3[is.na(Byphecode_mi3)]<-0
head(Byphecode_mi3)
lapply(Byphecode_mi3[c(5:7)], function(x) summary(as.factor(x)))

```


Number of phecode se, coloc and omim matches per drug-gene pair
```{r,echo=F}
 descriptivetable2druggene_mi=Byphecode_mi3%>% group_by(drugname,entrez_id) %>% summarise(total_mi=sum(PHECODE_MI), total_coloc=sum(coloc08_phecode), total_omim=sum(omim_phecode))

head(descriptivetable2druggene_mi)
descriptivetable2_mi=lapply(c('mi','coloc','omim'), function(x){
   column=paste0('total_',x)
   table=rbind.data.frame(x=cbind.data.frame(total=sum(descriptivetable2druggene_mi[[column]]),
                                                                           percentage=round(sum((descriptivetable2druggene_mi[[column]])/nrow(Byphecode_mi3)*100),2),
                                                                           Mean=round(mean(descriptivetable2druggene_mi[[column]]),2),
                                                                           Std=round(sd(descriptivetable2druggene_mi[[column]]),2)))
   rownames(table)=paste(x, 'phecode trait per DG pair')
      table                    
 })

descriptivetable2_mi<-do.call('rbind',descriptivetable2_mi)
descriptivetable2_mi
```

#### Chisq test
1.Phecode MI and OMIM phecode 
```{r, echo=F}
################################# chisq test all dg pairs ##################################################################
#omim
Dataset_mi_dg_omim=rbind.data.frame('Phecode mi yes'=(cbind('omim phecode Yes'=length(Byphecode_mi3$PHECODE_MI[Byphecode_mi3$PHECODE_MI==1 & Byphecode_mi3$omim_phecode==1]),'omim phecode No'= length(Byphecode_mi3$PHECODE_MI[Byphecode_mi3$PHECODE_MI==1 & Byphecode_mi3$omim_phecode==0]))),
'Phecode mi no'=(cbind('omim phecode Yes'=length(Byphecode_mi3$PHECODE_MI[Byphecode_mi3$PHECODE_MI==0 & Byphecode_mi3$omim_phecode==1]),'omim phecode No'= length(Byphecode_mi3$PHECODE_MI[Byphecode_mi3$PHECODE_MI==0 & Byphecode_mi3$omim_phecode==0]))))
Dataset_mi_dg_omim
D3_chisq_mi.omim=chisq.test(Dataset_mi_dg_omim,correct = FALSE)
D3_chisq_mi.omim
#'Warning message: In chisq.test(Dataset_mi_dg_omim, correct = FALSE)   Chi-squared approximation may be incorrect'
D3_chisq_mi.omim=chisq.test(Dataset_mi_dg_omim,correct = FALSE,simulate.p.value = TRUE)
D3_chisq_mi.omim
```

2.Phecode se and coloc phecode 
```{r,echo=F}
#coloc
Dataset_mi_dg_coloc=rbind.data.frame('Phecode mi yes'=(cbind('coloc phecode Yes'=length(Byphecode_mi3$PHECODE_MI[Byphecode_mi3$PHECODE_MI==1 & Byphecode_mi3$coloc08_phecode==1]),'coloc phecode No'= length(Byphecode_mi3$PHECODE_MI[Byphecode_mi3$PHECODE_MI==1 & Byphecode_mi3$coloc08_phecode==0]))),
                                    'Phecode mi no'=(cbind('coloc phecode Yes'=length(Byphecode_mi3$PHECODE_MI[Byphecode_mi3$PHECODE_MI==0 & Byphecode_mi3$coloc08_phecode==1]),'coloc phecode No'= length(Byphecode_mi3$PHECODE_MI[Byphecode_mi3$PHECODE_MI==0 & Byphecode_mi3$coloc08_phecode==0]))))

Dataset_mi_dg_coloc
D3_chisq_mi.coloc=chisq.test(Dataset_mi_dg_coloc,correct = FALSE)
D3_chisq_mi.coloc
```
 
#### Models 
 1.logistic model
```{r}
logmod_coloc_mi=glm(PHECODE_MI~coloc08_phecode, data = Byphecode_mi3, family = 'binomial')
summary(logmod_coloc_mi)

logmod_omim_mi=glm(PHECODE_MI~omim_phecode, data = Byphecode_mi3, family = 'binomial')
summary(logmod_omim_mi)
 
```
 
 2.mixed model
```{r,echo=F}
mi_predictbycoloc=glmer(PHECODE_MI~coloc08_phecode+(1|PheCode), data = Byphecode_mi3, family = 'binomial')
summary(mi_predictbycoloc) 

mi_predictbyomim=glmer(PHECODE_MI~omim_phecode+(1|PheCode), data = Byphecode_mi3, family = 'binomial')
summary(mi_predictbyomim)

multi_mi=glm(PHECODE_MI~coloc08_phecode+omim_phecode, data = Byphecode_mi3, family = 'binomial')
summary(multi_mi)

```



### Grouped MI Phecode child terms by lowest level
1. Dataset - 590 unique drugs. 2939 unique drug-gene pairs X 651 unique SE = 2,060,239 lines

``` {r, echo=F}
mi_dataset2=mi_dataset[c(-9,-11)]
mi_dataset2$PHECODE_MI_grouped1=ifelse(!is.na(str_extract(mi_dataset2$PHECODE_MI,'\\.[0-9]')), paste0(0,str_extract(mi_dataset2$PHECODE_MI,'\\.[0-9]')), 0)
mi_dataset2$PHECODE_MI_grouped=trunc(mi_dataset2$PHECODE_MI)+as.numeric(mi_dataset2$PHECODE_MI_grouped1)
head(mi_dataset2)
mi_dataset2$coloc08_phecode_grouped1=ifelse(!is.na(str_extract(mi_dataset2$coloc08_phecode,'\\.[0-9]')), paste0(0,str_extract(mi_dataset2$coloc08_phecode,'\\.[0-9]')), 0)
mi_dataset2$coloc08_phecode_grouped=trunc(mi_dataset2$coloc08_phecode)+as.numeric(mi_dataset2$coloc08_phecode_grouped1)

mi_dataset2$omim_phecode_grouped1=ifelse(!is.na(str_extract(mi_dataset2$omim_phecode,'\\.[0-9]')), paste0(0,str_extract(mi_dataset2$omim_phecode,'\\.[0-9]')), 0)
mi_dataset2$omim_phecode_grouped=trunc(mi_dataset2$omim_phecode)+as.numeric(mi_dataset2$omim_phecode_grouped1)

dim(mi_dataset2)
# 7258727      10
drug_info_mi=mi_dataset2 %>% distinct(drugname, entrez_id, symbol)
dim(drug_info_mi)
dataset_by_phecode_collapsed_mi=mi_dataset2[c(1:3,12,14,16)]
dataset_by_phecode_collapsed_mi=dataset_by_phecode_collapsed_mi[!duplicated(dataset_by_phecode_collapsed_mi),]
#0 if gene doesnt match phecode se. NA if omim or coloc never reported gene
#Dataset- 589 unique drugs. 2933 unique drug-gene pairs X 672 unique SE
mi_phecodes_collapsed=unique(dataset_by_phecode_collapsed_mi$PHECODE_MI_grouped)
 
Byphecode2_collapsed_mi<-lapply(c(mi_phecodes_collapsed), function(mi){
  #cat(paste(mi,'\n'))
  dataset_phecode=dataset_by_phecode_collapsed_mi[which(dataset_by_phecode_collapsed_mi$PHECODE_MI_grouped==mi),]
  drug_info_mi$PheCode=mi
  drug_info_one_all=merge(drug_info_mi,dataset_phecode, by= paste(colnames(drug_info_mi[-4])),all=T)
  drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 #cat(nrow(drug_info_one_all))
    drug_info_one_all[c(5:7)]<-lapply(drug_info_one_all[c(5:7)], function(col) ifelse(col==mi,1,0))
   drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 drug_info_one_al2=drug_info_one_all[order(drug_info_one_all$PHECODE_MI_grouped,drug_info_one_all$coloc08_phecode_grouped,drug_info_one_all$omim_phecode_grouped, decreasing=T),]
 drug_info_one_al2$PHECODE_MI_grouped[is.na(drug_info_one_al2$PHECODE_MI_grouped)]<-0
 drug_info_one_al2=drug_info_one_al2[!duplicated(drug_info_one_al2[c(1:4)]),]
  # cat(paste(nrow(drug_info_one_al2)),'\n')
 drug_info_one_al2=drug_info_one_al2
})
Byphecode2_collapsed_mi<-do.call('rbind',Byphecode2_collapsed_mi)  

Byphecode2_collapsed_mi[is.na(Byphecode2_collapsed_mi)]<-0
head(Byphecode2_collapsed_mi)
lapply(Byphecode2_collapsed_mi[c(5:7)], function(x) summary(as.factor(x)))

```

Number of phecode se, coloc and omim matches per drug-gene pair once child terms collapsed
```{r,echo=F}

 descriptivetable2druggene_collapsed_mi=Byphecode2_collapsed_mi%>% group_by(drugname,entrez_id) %>% summarise(total_mi=sum(PHECODE_MI_grouped), total_coloc=sum(coloc08_phecode_grouped), total_omim=sum(omim_phecode_grouped))

head(descriptivetable2druggene_collapsed_mi)
descriptivetable2_collapsed_mi=lapply(c('mi','coloc','omim'), function(x){
   column=paste0('total_',x)
   table=rbind.data.frame(x=cbind.data.frame(total=sum(descriptivetable2druggene_collapsed_mi[[column]]),
                                                                           percentage=round(sum((descriptivetable2druggene_collapsed_mi[[column]])/nrow(Byphecode2_collapsed_mi)*100),2),
                                                                           Mean=round(mean(descriptivetable2druggene_collapsed_mi[[column]]),2),
                                                                           Std=round(sd(descriptivetable2druggene_collapsed_mi[[column]]),2)))
   rownames(table)=paste(x, 'phecode trait per DG pair')
      table                    
 })

descriptivetable2_collapsed_mi<-do.call('rbind',descriptivetable2_collapsed_mi)
descriptivetable2_collapsed_mi

```

#### Chisq test
1.Phecode se and OMIM phecode
```{r,echo=F}
################################# chisq test all dg pairs ##################################################################
#omim
Dataset_dg_omim_collapse_mi=rbind.data.frame('Phecode mi yes'=(cbind('omim phecode Yes'=length(Byphecode2_collapsed_mi$PHECODE_MI_grouped[Byphecode2_collapsed_mi$PHECODE_MI_grouped==1 & Byphecode2_collapsed_mi$omim_phecode_grouped==1]),'omim phecode No'= length(Byphecode2_collapsed_mi$PHECODE_MI_grouped[Byphecode2_collapsed_mi$PHECODE_MI_grouped==1 & Byphecode2_collapsed_mi$omim_phecode_grouped==0]))),
'Phecode mi no'=(cbind('omim phecode Yes'=length(Byphecode2_collapsed_mi$PHECODE_MI_grouped[Byphecode2_collapsed_mi$PHECODE_MI_grouped==0 & Byphecode2_collapsed_mi$omim_phecode_grouped==1]),'omim phecode No'= length(Byphecode2_collapsed_mi$PHECODE_MI_grouped[Byphecode2_collapsed_mi$PHECODE_MI_grouped==0 & Byphecode2_collapsed_mi$omim_phecode_grouped==0]))))
Dataset_dg_omim_collapse_mi
D3_chisq_mi.omim_collapse=chisq.test(Dataset_dg_omim_collapse_mi,correct = FALSE)
D3_chisq_mi.omim_collapse
#'Warning message: In chisq.test(Datasetv3_dg_omim, correct = FALSE)   Chi-squared approximation may be incorrect'
D3_chisq_mi.omim_collapse=chisq.test(Dataset_dg_omim_collapse_mi,correct = FALSE,simulate.p.value = TRUE)
D3_chisq_mi.omim_collapse
```
1.Phecode se and coloc phecode
```{r,echo=F}
#coloc
Dataset_dg_coloc_collapse_mi=rbind.data.frame('Phecode mi yes'=(cbind('coloc phecode Yes'=length(Byphecode2_collapsed_mi$PHECODE_MI_grouped[Byphecode2_collapsed_mi$PHECODE_MI_grouped==1 & Byphecode2_collapsed_mi$coloc08_phecode_grouped==1]),'coloc phecode No'= length(Byphecode2_collapsed_mi$PHECODE_MI_grouped[Byphecode2_collapsed_mi$PHECODE_MI_grouped==1 & Byphecode2_collapsed_mi$coloc08_phecode_grouped==0]))),
                                    'Phecode se no'=(cbind('coloc phecode Yes'=length(Byphecode2_collapsed_mi$PHECODE_MI_grouped[Byphecode2_collapsed_mi$PHECODE_MI_grouped==0 & Byphecode2_collapsed_mi$coloc08_phecode_grouped==1]),'coloc phecode No'= length(Byphecode2_collapsed_mi$PHECODE_MI_grouped[Byphecode2_collapsed_mi$PHECODE_MI_grouped==0 & Byphecode2_collapsed_mi$coloc08_phecode_grouped==0]))))

Dataset_dg_coloc_collapse_mi
D3_chisq_mi.coloc_collapse=chisq.test(Dataset_dg_coloc_collapse_mi,correct = FALSE)
D3_chisq_mi.coloc_collapse
```
#### Models
1.logistic model 

```{r}
logmod_coloc_collapsed_mi=glm(PHECODE_MI_grouped~coloc08_phecode_grouped, data = Byphecode2_collapsed_mi, family = 'binomial')
summary(logmod_coloc_collapsed_mi)

logmod_omim_collapsed_mi=glm(PHECODE_MI_grouped~omim_phecode_grouped, data = Byphecode2_collapsed_mi, family = 'binomial')
summary(logmod_omim_collapsed_mi)
```

2.mixed model
```{r,echo=F}
mi_predictbycoloc_collapsed=glmer(PHECODE_MI_grouped~coloc08_phecode_grouped+(1|PheCode), data = Byphecode2_collapsed_mi, family = 'binomial')
summary(mi_predictbycoloc_collapsed) 

mi_predictbyomim_collapsed=glmer(PHECODE_MI_grouped~omim_phecode_grouped+(1|PheCode), data = Byphecode2_collapsed_mi, family = 'binomial')
summary(mi_predictbyomim_collapsed)

multi_collapsed_mi=glm(PHECODE_MI_grouped~coloc08_phecode_grouped+omim_phecode_grouped, data = Byphecode2_collapsed_mi, family = 'binomial')
summary(multi_collapsed_mi)
```

### Grouped Phecode child terms by lowest level. Genes collapsed
2. Dataset - 590 unique drugs X 651 unique SE = 384,090 lines
All genes aggregated per drug
``` {r, echo=F}

##group all genes togther 

#Dataset- 589 unique drugs X 672 unique SE
#395,808 lines  

Byphecode_agggregategene_collapsed_mi<-lapply(c(mi_phecodes_collapsed), function(mi){
 # cat(paste(se,'\n'))
  dataset_phecode=dataset_by_phecode_collapsed_mi[which(dataset_by_phecode_collapsed_mi$PHECODE_MI_grouped==mi),]
  drug_info_mi$PheCode=mi
  drug_info_one_all=merge(drug_info_mi,dataset_phecode, by= paste(colnames(drug_info_mi[-4])),all=T)
  drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
 #cat(nrow(drug_info_one_all))
    drug_info_one_all[c(5:7)]<-lapply(drug_info_one_all[c(5:7)], function(col) ifelse(col==mi,1,0))
    drug_info_one_all[is.na(drug_info_one_all)]<-0
   drug_info_one_all=drug_info_one_all[!duplicated(drug_info_one_all),]
   drug_info_one_all3=aggregate(.~drugname, drug_info_one_all[c(-2,-3,-4)], sum)
   drug_info_one_all3$Phecode=mi
   drug_info_one_all3[c(-1,-5)][drug_info_one_all3[c(-1,-5)]>1]<-1
  #  cat(paste(nrow(drug_info_one_all3)),'\n')
 drug_info_one_all3=drug_info_one_all3
})
Byphecode_agggregategene_collapsed_mi<-do.call('rbind',Byphecode_agggregategene_collapsed_mi)  
head(Byphecode_agggregategene_collapsed_mi)
lapply(Byphecode_agggregategene_collapsed_mi[c(2:4)], function(x) summary(as.factor(x)))

```
Number of unique phecode se, coloc and omim matches per drug once child terms collapsed
```{r,echo=F}

descriptivetable2druggene2_collapse_mi=Byphecode_agggregategene_collapsed_mi%>% group_by(drugname) %>% summarise(total_mi=sum(PHECODE_MI_grouped), total_coloc=sum(coloc08_phecode_grouped), total_omim=sum(omim_phecode_grouped))
head(descriptivetable2druggene2_collapse_mi)
 
 descriptivetable_perdrug_collapse_mi=lapply(c('mi','coloc','omim'), function(x){
   column=paste0('total_',x)
   table=rbind.data.frame(x=cbind.data.frame(total=sum(descriptivetable2druggene2_collapse_mi[[column]]),                                                                           percentage=round(sum((descriptivetable2druggene2_collapse_mi[[column]])/nrow(Byphecode_agggregategene_collapsed_mi)*100),2),
                                                                           Mean=round(mean(descriptivetable2druggene2_collapse_mi[[column]]),2),
                                                                           Std=round(sd(descriptivetable2druggene2_collapse_mi[[column]]),2)))
   rownames(table)=paste(x, 'phecode trait per drug')
      table                    
 })

descriptivetable_perdrug_collapse_mi<-do.call('rbind',descriptivetable_perdrug_collapse_mi)
descriptivetable_perdrug_collapse_mi
```

#### Chisq test
1.Phecode se and OMIM phecode
```{r,echo = F}
################################# chisq test aggregate dg pairs ##################################################################

#omim
Dataset_mi_dg_omim_aggregate_collapse=rbind.data.frame('Phecode mi yes'=(cbind('omim phecode Yes'=length(Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped[Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped==1 & Byphecode_agggregategene_collapsed_mi$omim_phecode_grouped==1]),'omim phecode No'= length(Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped[Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped==1 & Byphecode_agggregategene_collapsed_mi$omim_phecode_grouped==0]))),
                                             'Phecode mi no'=(cbind('omim phecode Yes'=length(Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped[Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped==0 & Byphecode_agggregategene_collapsed_mi$omim_phecode_grouped==1]),'omim phecode No'= length(Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped[Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped==0 & Byphecode_agggregategene_collapsed_mi$omim_phecode_grouped==0]))))

D3ag_chisq_mi.omim_collapse=chisq.test(Dataset_mi_dg_omim_aggregate_collapse,correct = FALSE)
D3ag_chisq_mi.omim_collapse
D3ag_chisq_mi.omim_collapse=chisq.test(Dataset_mi_dg_omim_aggregate_collapse,correct = FALSE,simulate.p.value = TRUE)
D3ag_chisq_mi.omim_collapse
```

2.Phecode se and coloc phecode
```{r,echo=F}
#coloc

Dataset_mi_dg_coloc_aggregate_collapse=rbind.data.frame('Phecode mi yes'=(cbind('coloc phecode Yes'=length(Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped[Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped==1 & Byphecode_agggregategene_collapsed_mi$coloc08_phecode_grouped==1]),'coloc phecode No'= length(Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped[Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped==1 & Byphecode_agggregategene_collapsed_mi$coloc08_phecode_grouped==0]))),
                                              'Phecode mi no'=(cbind('coloc phecode Yes'=length(Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped[Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped==0 & Byphecode_agggregategene_collapsed_mi$coloc08_phecode_grouped==1]),'coloc phecode No'= length(Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped[Byphecode_agggregategene_collapsed_mi$PHECODE_MI_grouped==0 & Byphecode_agggregategene_collapsed_mi$coloc08_phecode_grouped==0]))))

D3ag_chisq_mi.coloc_collapse=chisq.test(Dataset_mi_dg_coloc_aggregate_collapse,correct = FALSE)

```

#### Models
1.logistic model

```{r}

logmod_coloc_drug_collapse_mi=glm(PHECODE_MI_grouped~coloc08_phecode_grouped, data = Byphecode_agggregategene_collapsed_mi, family = 'binomial')
summary(logmod_coloc_drug_collapse_mi)
  
logmod_omim_drug_collapse_mi=glm(PHECODE_MI_grouped~omim_phecode_grouped, data = Byphecode_agggregategene_collapsed_mi, family = 'binomial')
summary(logmod_omim_drug_collapse_mi)

```

2.mixed model

```{r,echo=F}
mi_predictbycoloc_groupgene_collapse=glmer(PHECODE_MI_grouped~coloc08_phecode_grouped+(1|Phecode), family = 'binomial',data = Byphecode_agggregategene_collapsed_mi)
summary(mi_predictbycoloc_groupgene_collapse)
 
mi_predictbyomim_groupgene_collapse=glmer(PHECODE_MI_grouped~omim_phecode_grouped+(1|Phecode), family = 'binomial', data = Byphecode_agggregategene_collapsed_mi )
summary(mi_predictbyomim_groupgene_collapse)

#wflow_publish(c('analysis/Models_drugdataset.Rmd'), 'updating models')

```
