---
title: "Side effect info in SIDER 2"
author: "AineDuffy"
date: "2020-04-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
####run with sider 4. Only did sider 2 becuase no adr score. now if imputing them can do with s4


####Looking at the frequency of side effects and drugs in SIDER V2 (2012-10-17). Data sourced from FTP link from [Here](http://sideeffects.embl.de) 

```{r, include=F}

library(data.table)
library(stringr)
library(xlsx)
library(ggplot2) 
library(dplyr)
library(kableExtra)
library(DT)
library(scales)
```

### Overview of dataset 
``` {r, echo=F}
se=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/SIDER 2/meddra_adverse_effects.tsv.gz',data.table=F)
colnames(se)=c('stich1','stich2','umls_onlabel','drugname','se','medra_concepttype','umls_medra','medra_se')

print(paste('There are ',length(unique(se$drugname)), 'unique drugs with se' )) #
print(paste('There are ',length(unique(se$medra_se)), 'unique se' ))
se=se[which(se$medra_concepttype=='PT'),]
se=se %>% group_by(drugname)%>% mutate(se_no = row_number())
print(paste('There are ',length(unique(se$drugname)), 'unique drugs with PT se' ))
print(paste('There are ',length(unique(se$medra_se)), 'unique PT se' )) #There are  3209 unique PT se"
```
#### Taking just preferred term 'PT' side effects
```{r, echo=F}

#996 drugs 
#4492 UNIQUE SE. #3209 PT

se_table_a=se[c(4,7,8,9)][!duplicated(se[c(4,7,8)]),] #103156
se_table_a$se_no=gsub('$',':medra',se_table_a$se_no)
colnames(se_table_a)[c(2,3)]=c('umls','se')
se_table_b=se[c(4,3,5,9)][!duplicated(se[c(4,3,5)]),] #99423
se_table_b$se_no=gsub('$',':label',se_table_b$se_no)

colnames(se_table_b)[c(2,3)]=c('umls','se')
se_table_1=rbind(se_table_a,se_table_b)
se_table_1$se=str_to_lower(se_table_1$se)
se_table_1=se_table_1[!duplicated(se_table_1[c(1,2,3)]),]#147217
dim(se_table_1)

se_tab_1=as.data.frame(se_table_1 %>% select(drugname, se) %>%
 group_by(se) %>% tally() %>%
  arrange(desc(n)))

###Side effect frequency
DT::datatable(se_tab_1,caption = 'Table 1: Side effect frequency in Sider', filter='top',colnames=c('Side effects','Frequency'), editable = TRUE, extensions = 'Buttons', options = list(dom='Bfrtip', buttons='copy','csvHtml15','excel'))

####Side effect frequency summary
summary(se_tab_1$n)

ggplot(se_tab_1, aes(x=n))+ geom_histogram(binwidth = 20, color="black", fill="light blue") +
  labs(x='Side effect frequency',y='Number of drugs', title='Side effect frequency across SIDER2') +
  theme_classic()

```

### Main Indications

```{r, echo =F}
mi=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/SIDER 2/indications_raw.tsv.gz', data.table=F, header=F)
colnames(mi)=c('label_identifier','cui','mi')

label=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/SIDER 2/label_mapping.tsv', data.table=F)
colnames(label)=c('drugname','brandrugname','markermapped','stich_id','stich_id1','url','label_identifier')

label1=as.data.frame(do.call(rbind, apply(label[c(2,7)],1,function(X){
	do.call(expand.grid, strsplit(X, ';',useBytes = TRUE))
}
	)))

label2=merge(label1,label[c(1,7)], by='label_identifier',all=T)
label3=label2[!duplicated(label2),]
label3$brandrugname=str_to_lower(label3$brandrugname)
label3$drugname=str_to_lower(label3$drugname)
colnames(label3)=c('label_identifier','drugname','drugname')
label4=rbind(label3[c(1,2)], label3[c(1,3)])
label4=label4[!duplicated(label4),] #72258 rows

mi_label=inner_join(mi, label4, by='label_identifier')
mi_label$mi=str_to_lower(mi_label$mi)
mi_label1=mi_label[-1]
mi_label1=mi_label1[!duplicated(mi_label1),]
colnames(mi_label1)[1]='cui_mi'
#65573 line
paste(length(unique(mi_label1$drugname)),'unique drugs with', length(unique(mi_label1$cui)), 'unique main indications')

#Drug frequency for main indications
mitab=as.data.frame(mi_label1 %>% 
 group_by(mi) %>% tally() %>%
  arrange(desc(n)))
DT::datatable(mitab,caption = 'Table 2: Main indication frequency in Sider', filter='top',colnames=c('Main indication','Frequency'), editable = TRUE, extensions = 'Buttons', options = list(dom='Bfrtip', buttons='copy','csvHtml15','excel'))

summary(mitab$n)

ggplot(mitab, aes(x=n))+ geom_histogram(binwidth = 20, color="black", fill="light blue") +
  labs(x='Main indication frequency',y='Number of drugs', title='Main indication frequency across SIDER2') +
  theme_classic()




```
### Filter SE.
##### 1. Remove se for drugs where se==mi. Used both umls label and medra se cui 
##### 2. How many side effects are there that are observed in >=10% of drugs? How many side effects in total?
##### 3. How many drugs are there that are observed in >=10% of side effect? How many drugs in total?
```{r, echo=F}

#1) Remove se for drugs ==mi. used both umls label and medra se
drugs_se_main_1=left_join(se_table_1,mi_label1, by='drugname')

#784 drugs which have mi and se

paste(length(unique(drugs_se_main_1$drugname)), 'drugs have se and',length(unique(drugs_se_main_1$drugname[!is.na(drugs_se_main_1$mi)])),'of these have mi info as well.', length(unique(drugs_se_main_1$se)), 'unique se and',length(unique(drugs_se_main_1$mi)), 'unique mi' )
#drugs_se_main=drugs_se_main[!duplicated(drugs_se_main),]
matchse_mi_1=drugs_se_main_1[which(drugs_se_main_1$cui_mi==drugs_se_main_1$umls),]
head(matchse_mi_1)
paste(length(unique(matchse_mi_1$drugname)), 'drugs atleast one exact cui match for mi = se')


#matches removed
remove.matchse_mi_1=drugs_se_main_1[which(drugs_se_main_1$cui_mi!=drugs_se_main_1$umls|is.na(drugs_se_main_1$mi)),]

se_table_matchmiremoved_1=remove.matchse_mi_1[c(1:4)][!duplicated(remove.matchse_mi_1[c(1:4)]),]
setab_1=as.data.frame(se_table_matchmiremoved_1 %>% select(drugname, se) %>% distinct() %>%
 group_by(se) %>% tally() %>%
  arrange(desc(n)))

DT::datatable(setab_1,caption = 'Table 3: Side effect frequency in Sider for drugs with main indication info', filter='top',colnames=c('Side effects','Frequency'), editable = TRUE, extensions = 'Buttons', options = list(dom='Bfrtip', buttons='copy','csvHtml15','excel'))

#1) How many side effects are there that are observed in >=10% of drugs? How many side effects in total?
se.greater10_1=setab_1[setab_1$n>length(unique(se_table_matchmiremoved_1$drugname))/10,]
print(paste(nrow(se.greater10_1),'out of',length(unique(se_table_matchmiremoved_1$se)),'unique side effects observed in > 10% of',length(unique(se_table_matchmiremoved_1$drugname)),'drugs')) 

#2) How many drugs are there that are observed in >=10% of side effect? How many drugs in total?
#drugs with more than >10% side effects?
setabdrug_1=as.data.frame(se_table_matchmiremoved_1 %>% 
 group_by(drugname) %>% tally() %>%
  arrange(desc(n)))

drug.greater10_1=setabdrug_1[setabdrug_1$n>length(unique(se_table_matchmiremoved_1$se))/10,]
print(paste(nrow(drug.greater10_1),'out of',length(unique(se_table_matchmiremoved_1$drugname)), 'drugs observed with > 10% of',length(unique(se_table_matchmiremoved_1$se)),'unique se'))

filterdrugs_se_1=subset(remove.matchse_mi_1, !(drugname %in% drug.greater10_1$drugname))
filterdrugs_se_1=subset(filterdrugs_se_1, !(se %in% se.greater10_1$se))
paste(length(unique(filterdrugs_se_1$drugname)),'unique drugs and', length(unique(filterdrugs_se_1$se)),'unique se once removed >10% se and drugs')


```

### Match side effects to ICD10 codes 
```{r, echo=F}
###Side effects matched to ICD10 codes
icd_cui=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/MRCONSO.RRF.txt',data.table=F, quote='')
icd_cui=icd_cui %>% filter(V12=='ICD10CM',V2=='ENG' ) %>% distinct(V1,V11) %>% rename(umls=V1,ICD10=V11)
#matched icd codes to both label and medra

se_icd_all_1=merge(filterdrugs_se_1, icd_cui, by='umls')
colnames(se_icd_all_1)=gsub('ICD10','ICD10_SE', colnames(se_icd_all_1))
dim(se_icd_all_1)
se_icd_all_1=se_icd_all_1[!duplicated(se_icd_all_1[-4]),]
dim(se_icd_all_1)
print(paste('There are',length(unique(se_icd_all_1$drugname)), 'unique drugs with se that match to ICD10' )) #872 unique drugs
print(paste('There are',length(unique(se_icd_all_1$se)), 'unique se' ))
print(paste('There are',length(unique(se_icd_all_1$ICD10_SE)), 'unique ICD codes' ))



``` 

#### some se have mutliple icd10 codes. remove after match to phecode
```{r, echo=F}

se_1=as.data.frame(se_icd_all_1 %>% select(se, ICD10_SE) %>% distinct() %>% group_by(se) %>% filter(n()>1))
head(se_1)

###frequency of each side effect 
setab_icd_1=se_icd_all_1 %>% select(drugname, se,ICD10_SE) %>% distinct() %>% 
 group_by(se,ICD10_SE)  %>% distinct() %>% tally() %>%
  arrange(desc(n))
print(paste('There are',length(unique(se_1$se)), 'unique medra_se with duplicate ICD10 codes' ))

 
###Side effect frequency for each drug once match to icd
DT::datatable(setab_icd_1, caption = 'Table 4: Side effect frequency in Sider matched to ICD10 code',filter='top', colnames=c('Side effects','ICD10'  ,'Frequency'), editable = TRUE, extensions = 'Buttons', options = list(dom='Bfrtip', buttons='copy'))

####Side effect frequency summary

summary(setab_icd_1$n)
ggplot(setab_icd_1, aes(x=n))+ geom_histogram(binwidth = 5, color="black", fill="light blue") +
  labs(x='Side effect frequency',y='Frequency', title='Side effect frequency which match to ICD10 codes') +
  theme_classic()




```

#### Frequency of se per drugs

```{r, echo=F}
drugse_1=se_icd_all_1 %>% select(drugname, se) %>% distinct() %>%  group_by(drugname) %>% tally() %>%
  arrange(desc(n))
summary(drugse_1$n)
ggplot(drugse_1, aes(x=n))+ geom_histogram(bins=20,color="black", fill="light blue") +
  labs(x='Number of Side effects per drug',y='Number of drugs', title='Number of Side effects per drug') +
  theme_classic()

```

### ADR risk score
```{r, echo=F}

adr=read.xlsx('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/ADR_rankscore.xlsx', sheetIndex=1)
adr$Name=str_to_lower(adr$Name)

se_icd_all_1$se_adr=adr$Rank.score[match(se_icd_all_1$se, adr$Name)]
dim(se_icd_all_1)
#868 drugs which have se- icd code
paste(length(unique(se_icd_all_1$se[!is.na(se_icd_all_1$se_adr)])), 'out of', length(unique(se_icd_all_1$se)), 'unique se have a adr score')
"1053 out of 1830 unique se have a adr score"
drugadr_1=se_icd_all_1  %>% select(drugname, se, se_adr) %>% distinct()

drug1_1=drugadr_1 %>% group_by(drugname) %>% summarise(total.no_adr=n(), meanadr=mean(se_adr, na.rm=T), sd=sd(se_adr, na.rm = T)) %>%arrange(desc(total.no_adr))
DT::datatable(drug1_1)
summary(drug1_1[-1])

adr_se_1=se_icd_all_1 %>% select(se,se_adr) %>% distinct()

ggplot(adr_se_1, aes(x=se_adr))+ geom_histogram(bins = 10, color="black", fill="light blue") +
  labs(x='ADR',y='Side effect', title='Distrubution of ADR scores') +
  theme_classic()

adr_se_1$freq=setab_1$n[match(adr_se_1$se, setab_1$se)]
adr_se_1=adr_se_1[!is.na(adr_se_1$se_adr),]  

#proportion of drugs, divide by total no of drugs (784)
adr_se_1$freq=adr_se_1$freq/length(unique(drugadr_1$drugname))
cor_sp_1=cor.test(adr_se_1$se_adr, (adr_se_1$freq), method = 'spearman', exact=F)
cor_sp_1

ggplot(adr_se_1, aes(x=se_adr,y=freq))+geom_point()+ geom_smooth() +labs(x='ADR score', y='Se freq', title = 'Correlation between adr score and se frequency') + annotate("text",color='red', x = .86, y = .725, label = paste('spearman coefficient =', round(cor_sp_1$estimate,4), '\n', 'P value = ',  scientific(cor_sp_1$p.value, digits = 3))) +  theme_classic() 

#generalized additive mode smoothing. #used spearmans correlation becuase side effect freq not normally distrubted

```

### Match ICD to Phecodes and get ADR phecode score

```{r, echo=F}

#ICDCODES TO SAIGE
ICD=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/ICD10CM2PHECODE.csv',data.table=F)
icd_adr=se_icd_all_1 %>% select(ICD10_SE, se_adr) %>% distinct() %>%filter(!is.na(se_adr))

#some duplicate scores for icd codes with dup se. map the se adr risk scores to phecodes
ICD$ADR=icd_adr$se_adr[match(ICD$ICD10CM, icd_adr$ICD10_SE)]


#phecode to icd frequency 
##each icd maps to only one phecode but multipe icds map to same 

se_icd_all_1$PHECODE_SE=ICD$PHECODE[match(se_icd_all_1$ICD10_SE,ICD$ICD10CM)]
dim(se_icd_all_1)
se_icd_all_1$PHECODE_DESC_SE=ICD$PHECODE_DESC[match(se_icd_all_1$ICD10_SE,ICD$ICD10CM)]
dim(se_icd_all_1)
se_icd_all_1=se_icd_all_1[!duplicated(se_icd_all_1[-4]),]
dim(se_icd_all_1)
se_icd_all_1=se_icd_all_1[!is.na(se_icd_all_1$PHECODE_SE),] 
Phenotype_cateogory=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/phecode_definitions1.2.csv',data.table=F)
se_icd_all_1$phecode_category=Phenotype_cateogory$category[match(se_icd_all_1$PHECODE_SE, Phenotype_cateogory$phecode)]
#Map number of distinct se + their phecodes to the phecode cateogory
Phecodefreq=as.data.frame(se_icd_all_1 %>% select(PHECODE_SE,phecode_category) %>% distinct() %>% group_by(phecode_category)   %>% tally(name = 'number_phecodes'))
se_phecodecat=as.data.frame(se_icd_all_1 %>% select(se,phecode_category) %>% distinct() %>% group_by(phecode_category) %>% tally(name = 'number_se'))
Phefre=merge(Phecodefreq, se_phecodecat,by='phecode_category')
Phefre1=reshape2::melt(Phefre,id.var='phecode_category')                         

ggplot(Phefre1, aes(x=phecode_category, y=value, fill=variable))+ geom_bar(stat = "identity",position='dodge') +theme_classic()+theme(axis.text.x = element_text(angle=-90)) + labs(x='Phecode cateogories', y='Number of side effects') 

```
### Main Indications to ICD

``` {r, echo=F}

colnames(icd_cui)[1]='cui_mi'
mi_label1$ICD10_MI=icd_cui$ICD10[match(mi_label1$cui_mi, icd_cui$cui_mi)]
paste(length(unique(mi_label1$mi[!is.na(mi_label1$ICD10_MI)])), 'mi have ICD10 code')
#1957 MI UNIQUE WITH ICD CODE FROM 2669

mi_label1$PHECODE_MI=ICD$PHECODE[match(mi_label1$ICD10_MI,ICD$ICD10CM)] #11 ICD codes didnt match to phecode
mi_label1$PHECODE_DESC_MI=ICD$PHECODE_DESC[match(mi_label1$ICD10_MI, ICD$ICD10CM)]

se_icd_all_1$ICD10_MI=mi_label1$ICD10_MI[match(se_icd_all_1$cui_mi,mi_label1$cui_mi)]
se_icd_all_1$PHECODE_MI=mi_label1$PHECODE_MI[match(se_icd_all_1$cui_mi,mi_label1$cui_mi)]


#MI AND SE which match on phecode 
matchingphecode_mi_se= se_icd_all_1 %>% select(se,se_no,ICD10_SE, PHECODE_SE, mi, ICD10_MI, PHECODE_MI) %>% filter(PHECODE_SE==PHECODE_MI) %>% distinct()

DT::datatable(matchingphecode_mi_se)

remove.matchse_miphecode=se_icd_all_1[which(se_icd_all_1$PHECODE_SE!=se_icd_all_1$PHECODE_MI|is.na(se_icd_all_1$PHECODE_MI)),]

#ADR SCORE THAT WAS MAPPED TO THE INITIAL SE. 
finalse_df= se_icd_all_1 %>% select(drugname,PHECODE_SE, se_adr,se_no) %>% distinct()
paste('Final datset is', nrow(finalse_df), 'rows,', length(unique(finalse_df$drugname)), 'drugs and', length(unique(finalse_df$PHECODE_SE)),'phecode se where',length(unique(finalse_df$PHECODE_SE[!is.na(finalse_df$se_adr)])),'phecodes have a matching se adr score')
library(tidyr)

finalse_df1=finalse_df[order(finalse_df$drugname,finalse_df$PHECODE_SE,finalse_df$se_adr),]
dim(finalse_df1)
finalse_df1=finalse_df1[!duplicated(finalse_df1[c(1:3)]),]
dim(finalse_df1)
finalse_df1=separate(finalse_df1,col=se_no, into=c('no','source'), sep=':',remove=F)
dim(finalse_df1)
finalse_df1=finalse_df1[!duplicated(finalse_df1[c(1,2,5)]),]
dim(finalse_df1)
finalse_df2<-finalse_df1[!(duplicated(finalse_df1[c(1,2)]) & is.na(finalse_df1$se_adr)),]
paste('Final datset is', nrow(finalse_df2), 'rows,', length(unique(finalse_df2$drugname)), 'drugs and', length(unique(finalse_df2$PHECODE_SE)),'phecode se where',length(unique(finalse_df2$PHECODE_SE[!is.na(finalse_df2$se_adr)])),'phecodes have a matching se adr score')
#dif adr scores for same phecode - 126 dif phecodes. Take max?
 DUPS=finalse_df2 %>% group_by(drugname,PHECODE_SE) %>% 
  filter(n()>1)
```

##MATCH FINAL DATASET TO GENES, GET TAU, TPM, COLOC RESULTS
### Match to entrez id  

```{r, echo=F}

drugbank=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/Target Drug-UniProt Links_ALL.csv',data.table=F, select=c(2,4))
#drugs and uniprot id
drugbank$Name=str_to_lower(drugbank$Name)
colnames(drugbank)[1]='drugname'
GeneDic <- fread("/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/GeneDictionary.txt", data.table = FALSE, select=c(2,19,26))
drugbank$entrez_id=GeneDic$entrez_id[match(drugbank$`UniProt ID`, GeneDic$uniprot_ids)]
drugbank$symbol=GeneDic$symbol[match(drugbank$`UniProt ID`, GeneDic$uniprot_ids)]
drugbank1=drugbank[!is.na(drugbank$entrez_id),]


finalse_df_gene=inner_join(drugbank[-2],finalse_df, by='drugname') #file -412794 rows
finalse_df_gene=finalse_df_gene[!is.na(finalse_df_gene$entrez_id),]
# #589 drugs #673 phecodes #658 genes

print(paste('There are',length(unique(finalse_df_gene$drugname)), 'unique drugs'))
print(paste('There are',length(unique(finalse_df_gene$entrez_id)), 'unique genes'))
print(paste('There are',length(unique(finalse_df_gene$PHECODE_SE)), 'unique phecodes and out of these', length(unique(finalse_df_gene$PHECODE_SE[!is.na(finalse_df_gene$se_adr)])), 'has an adr score'))

druggenes=finalse_df_gene %>% select(drugname, entrez_id) %>% group_by(drugname) %>% distinct()%>% tally() %>% arrange(desc(n))

ggplot(druggenes, aes(x=n))+ geom_histogram(binwidth =2, color="black",fill="light blue") +
  labs(x='Number of genes per drug',y='Number of drugs', title='Number of genes per drug') +
  theme_classic()
```

