---
title: "Side effect info in SIDER 2"
author: "AineDuffy"
date: "2020-04-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


####Looking at the frequency of side effects and drugs in SIDER V2 (2012-10-17). Data sourced from FTP link from [Here](http://sideeffects.embl.de) 

```{r, include=F}

library(data.table)
library(stringr)
library(xlsx)
library(ggplot2) 
library(dplyr)
library(kableExtra)
library(DT)
library(scales)
```

### Overview of dataset 
``` {r, echo=F}
se=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/SIDER 2/meddra_adverse_effects.tsv.gz',data.table=F)
colnames(se)=c('stich1','stich2','umls_onlabel','drugname','se','medra_concepttype','umls_medra','medra_se')
head(se)
print(paste('There are ',length(unique(se$drugname)), 'unique drugs with se' ))
print(paste('There are ',length(unique(se$medra_se)), 'unique se' ))
se=se[which(se$medra_concepttype=='PT'),]
print(paste('There are ',length(unique(se$drugname)), 'unique drugs with PT se' ))
print(paste('There are ',length(unique(se$medra_se)), 'unique PT se' ))
```
#### Taking just preferred term 'PT' side effects
```{r, echo=F}

#996 drugs 
#4492 UNIQUE SE. #3209 PT
se_table=se[c(4,7,8)][!duplicated(se[c(4,7,8)]),]
setab=as.data.frame(se_table %>% select(drugname, medra_se) %>%
 group_by(medra_se) %>% tally() %>%
  arrange(desc(n)))

###Side effect frequency
DT::datatable(setab,caption = 'Table 1: Side effect frequency in Sider', filter='top',colnames=c('Side effects','Frequency'), editable = TRUE, extensions = 'Buttons', options = list(dom='Bfrtip', buttons='copy','csvHtml15','excel'))

####Side effect frequency summary
summary(setab$n)

ggplot(setab, aes(x=n))+ geom_histogram(binwidth = 20, color="black", fill="light blue") +
  labs(x='Side effect frequency',y='Number of drugs', title='Side effect frequency across SIDER2') +
  theme_classic()

```

### Main Indications

```{r, echo =F}
mi=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/SIDER 2/indications_raw.tsv.gz', data.table=F, header=F)
colnames(mi)=c('label_identifier','cui','mi')

label=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/SIDER 2/label_mapping.tsv', data.table=F)
colnames(label)=c('drugname','brandrugname','markermapped','stich_id','stich_id1','url','label_identifier')

label1=as.data.frame(do.call(rbind, apply(label[c(2,7)],1,function(X){
	do.call(expand.grid, strsplit(X, ';',useBytes = TRUE))
}
	)))

label2=merge(label1,label[c(1,7)], by='label_identifier',all=T)
label3=label2[!duplicated(label2),]
label3$brandrugname=str_to_lower(label3$brandrugname)
label3$drugname=str_to_lower(label3$drugname)
colnames(label3)=c('label_identifier','drugname','drugname')
label4=rbind(label3[c(1,2)], label3[c(1,3)])
label4=label4[!duplicated(label4),] #72258 rows

mi_label=inner_join(mi, label4, by='label_identifier')
mi_label$mi=str_to_lower(mi_label$mi)
mi_label1=mi_label[-1]
mi_label1=mi_label1[!duplicated(mi_label1),]
colnames(mi_label1)[1]='cui_mi'
#65573 line
paste(length(unique(mi_label1$drugname)),'unique drugs with', length(unique(mi_label1$cui)), 'unique main indications')

#Drug frequency for main indications
mitab=as.data.frame(mi_label1 %>% 
 group_by(mi) %>% tally() %>%
  arrange(desc(n)))
DT::datatable(mitab,caption = 'Table 2: Main indication frequency in Sider', filter='top',colnames=c('Main indication','Frequency'), editable = TRUE, extensions = 'Buttons', options = list(dom='Bfrtip', buttons='copy','csvHtml15','excel'))

summary(mitab$n)

ggplot(mitab, aes(x=n))+ geom_histogram(binwidth = 20, color="black", fill="light blue") +
  labs(x='Main indication frequency',y='Number of drugs', title='Main indication frequency across SIDER2') +
  theme_classic()




```
### Filter SE
#### 1. Remove se for drugs where se==mi. Used both umls label and medra se cui 
#### 2. How many side effects are there that are observed in >=10% of drugs? How many side effects in total?
#### 3. How many drugs are there that are observed in >=10% of side effect? How many drugs in total?
```{r, echo=F}

#1) Remove se for drugs ==mi. used both umls label and medra se
drugs_se_main=inner_join(se[c(-1,-2,-6)],mi_label1, by='drugname')
#784 drugs which have mi and se
paste(length(unique(drugs_se_main$drugname)), 'drugs have se and mi info.', length(unique(drugs_se_main$medra_se)), 'unique se and',length(unique(drugs_se_main$mi)), 'unique mi' )
#drugs_se_main=drugs_se_main[!duplicated(drugs_se_main),]
matchse_mi1=drugs_se_main[which(drugs_se_main$cui_mi==drugs_se_main$umls_medra|drugs_se_main$cui_mi==drugs_se_main$umls_onlabel),]
head(matchse_mi1)
paste(length(unique(matchse_mi1$drugname)), 'drugs with exact cui match for mi = se')


#matches removed
remove.matchse_mi=drugs_se_main[which(drugs_se_main$cui_mi!=drugs_se_main$umls_medra|drugs_se_main$cui_mi!=drugs_se_main$umls_onlabel),]

se_table_matchmiremoved=remove.matchse_mi[c(1:5)][!duplicated(remove.matchse_mi[c(1:5)]),]
setab=as.data.frame(se_table_matchmiremoved %>% select(drugname, medra_se) %>% distinct() %>%
 group_by(medra_se) %>% tally() %>%
  arrange(desc(n)))

DT::datatable(setab,caption = 'Table 3: Side effect frequency in Sider for drugs with main indication info', filter='top',colnames=c('Side effects','Frequency'), editable = TRUE, extensions = 'Buttons', options = list(dom='Bfrtip', buttons='copy','csvHtml15','excel'))

#1) How many side effects are there that are observed in >=10% of drugs? How many side effects in total?
se.greater10=setab[setab$n>length(unique(se_table_matchmiremoved$drugname))/10,]
print(paste(nrow(se.greater10),'out of',length(unique(se_table_matchmiremoved$medra_se)),'unique side effects observed in > 10% of',length(unique(se_table_matchmiremoved$drugname)),'drugs')) 

#2) How many drugs are there that are observed in >=10% of side effect? How many drugs in total?
#drugs.wse.greater10=subset(se_table_matchmiremoved, (medra_se %in% se.greater10$medra_se))
#paste(length(unique(drugs.wse.greater10$drugname)), 'out of', length(unique(se_table_matchmiremoved$drugname)), 'drugs had a se that had > 10% frequency')

#drugs with more than >10% side effects?
setabdrug=as.data.frame(se_table_matchmiremoved %>% 
 group_by(drugname) %>% tally() %>%
  arrange(desc(n)))

drug.greater10=setabdrug[setabdrug$n>length(unique(se_table_matchmiremoved$medra_se))/10,]
print(paste(nrow(drug.greater10),'out of',length(unique(se_table_matchmiremoved$drugname)), 'drugs observed with > 10% of',length(unique(se_table_matchmiremoved$medra_se)),' unique se'))


```

### Match side effects to ICD10 codes 
```{r, echo=F}
###Side effects matched to ICD10 codes
icd_cui=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/MRCONSO.RRF.txt',data.table=F, quote='')
icd_cui=icd_cui %>% filter(V12=='ICD10CM',V2=='ENG' ) %>% distinct(V1,V11) %>% rename(umls_medra=V1,ICD10=V11)
se_icd=merge(se_table_matchmiremoved, icd_cui, by='umls_medra')
colnames(se_icd)=gsub('ICD10','ICD10_SE', colnames(se_icd))

print(paste('There are ',length(unique(se_icd$drugname)), 'unique drugs with se that match to ICD10' ))
print(paste('There are ',length(unique(se_icd$medra_se)), 'unique se' ))
print(paste('There are ',length(unique(se_icd$ICD10_SE)), 'unique ICD codes' ))
``` 

#### some se have mutliple icd10 codes
```{r, echo=F}

se1=as.data.frame(se_icd %>% select(medra_se, ICD10_SE) %>% distinct() %>% group_by(medra_se) %>% filter(n()>1))
head(se1)

###frequency of each side effect 
setab_icd=se_icd %>% select(drugname, medra_se,ICD10_SE) %>% distinct() %>% 
 group_by(medra_se,ICD10_SE) %>% tally() %>%
  arrange(desc(n))
print(paste('There are ',length(unique(se1$medra_se)), 'unique medra_se with duplicate ICD10 codes' ))

 
###Side effect frequency for each drug once match to icd
DT::datatable(setab_icd, caption = 'Table 4: Side effect frequency in Sider matched to ICD10 code',filter='top', colnames=c('Side effects','ICD10'  ,'Frequency'), editable = TRUE, extensions = 'Buttons', options = list(dom='Bfrtip', buttons='copy'))

####Side effect frequency summary

summary(setab_icd$n)
ggplot(setab_icd, aes(x=n))+ geom_histogram(binwidth = 20, color="black", fill="light blue") +
  labs(x='Side effect frequency',y='Frequency', title='Side effect frequency which match to ICD10 codes') +
  theme_classic()
```

### Filter SE after match to ICD10 codes. 
#### Less SE 
#### 1. How many side effects are there that are observed in >=10% of drugs? How many side effects in total?
#### 2. How many drugs are there that are observed in >=10% of side effect? How many drugs in total?
```{r, echo=F}
#1) How many side effects are there that are observed in >=10% of drugs? How many side effects in total?
se_icd.greater10=setab_icd[setab_icd$n>length(unique(se_icd$drugname))/10,]
print(paste(nrow(se_icd.greater10),'out of',length(unique(se_icd$medra_se)),'unique side effects observed in > 10% of',length(unique(se_icd$drugname)),'drugs')) 

#2) How many drugs are there that are observed in >=10% of side effect? How many drugs in total?
#drugs_icd.wse.greater10=subset(se_icd, (medra_se %in% se_icd.greater10$medra_se))
#paste(length(unique(drugs_icd.wse.greater10$drugname)), 'out of', length(unique(se_icd$drugname)), 'drugs had a se that had > 10% frequency')

#drugs with more than >10% side effects?
seicdtabdrug=as.data.frame(se_icd %>% select(drugname, medra_se) %>%
 group_by(drugname) %>% tally() %>%
  arrange(desc(n)))

drugicd.greater10=seicdtabdrug[seicdtabdrug$n>length(unique(se_icd$medra_se))/10,]
print(paste(nrow(drugicd.greater10),'out of',length(unique(se_icd$drugname)), 'drugs observed with > 10% of',length(unique(se_icd$medra_se)),' unique se'))


```


#### Frequency of se per drugs

```{r, echo=F}
drugse=se_icd %>% select(drugname, medra_se) %>% distinct() %>%  group_by(drugname) %>% tally() %>%
  arrange(desc(n))
summary(drugse$n)
ggplot(drugse, aes(x=n))+ geom_histogram(binwidth = 20, color="black", fill="light blue") +
  labs(x='Number of Side effects per drug',y='Number of drugs', title='Number of Side effects per drug') +
  theme_classic()

```

### ADR risk score
```{r, echo=F}

adr=read.xlsx('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/ADR_rankscore.xlsx', sheetIndex=1)
adr$Name=str_to_lower(adr$Name)

se_icd$medra_se=str_to_lower(se_icd$medra_se)
se_icd$se=str_to_lower(se_icd$se)

se_icd$se_adr1=adr$Rank.score[match(se_icd$medra_se, adr$Name)]
se_icd$se_adr2=adr$Rank.score[match(se_icd$se, adr$Name)]
se_icd$se_adr<-ifelse(!is.na(se_icd$se_adr1), se_icd$se_adr1, se_icd$se_adr2)
se_icd=se_icd %>% select(-se_adr2, -se_adr1)

paste(length(unique(se_icd$medra_se[!is.na(se_icd$se_adr)])), 'out of', length(unique(se_icd$medra_se)), 'unique se have a adr score')
##"1205 out of 1276 unique se have a adr score"
drugadr=se_icd  %>% select(drugname, medra_se, se_adr) %>% distinct()

drug1=drugadr %>% group_by(drugname) %>% summarise(total.no_adr=n(), meanadr=mean(se_adr, na.rm=T), sd=sd(se_adr, na.rm = T))
DT::datatable(drug1)
summary(drug1[-1])

adr_se=se_icd %>% select(medra_se,se_adr) %>% distinct()

ggplot(adr_se, aes(x=se_adr))+ geom_histogram(bins = 10, color="black", fill="light blue") +
  labs(x='ADR',y='Side effect', title='Distrubution of ADR scores') +
  theme_classic()

setab$medra_se=str_to_lower(setab$medra_se)
adr_se$freq=setab$n[match(adr_se$medra_se, setab$medra_se)]
adr_se=adr_se[!is.na(adr_se$se_adr),]  

#proportion of drugs, divide by total no of drugs (784)
adr_se$freq=adr_se$freq/length(unique(drugadr$drugname))
cor_sp=cor.test(adr_se$se_adr, (adr_se$freq), method = 'spearman', exact=F)
cor_sp
ggplot(adr_se, aes(x=se_adr,y=freq))+geom_point()+ geom_smooth() +labs(x='ADR score', y='Se freq', title = 'Correlation between adr score and se frequency') + annotate("text",color='red', x = .86, y = .725, label = paste('spearman coefficient =', round(cor_sp$estimate,4), '\n', 'P value = ',  scientific(cor_sp$p.value, digits = 3))) +  theme_classic() 
#generalized additive mode smoothing. #used spearmans correlation becuase side effect freq not normally distrubted

```

### Match ICD to Phecodes and get ADR phecode score

```{r, echo=F}

#ICDCODES TO SAIGE
ICD=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/ICD10CM2PHECODE.csv',data.table=F)
icd_adr=se_icd %>% select(ICD10_SE, se_adr) %>% distinct()

#some duplicate scores for icd codes with dup se 
ICD$ADR=icd_adr$se_adr[match(ICD$ICD10CM, icd_adr$ICD10_SE)]


#phecode to icd frequency 
##each icd maps to only one phecode but multipe icds map to same 

se_icd$PHECODE_SE=ICD$PHECODE[match(se_icd$ICD10_SE,ICD$ICD10CM)]
se_icd$PHECODE_DESC_SE=ICD$PHECODE_DESC[match(se_icd$ICD10_SE,ICD$ICD10CM)]
se_icd=se_icd[!duplicated(se_icd),]
 
Phenotype_cateogory=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/phecode_definitions1.2.csv',data.table=F)
se_icd$phecode_category=Phenotype_cateogory$category[match(se_icd$PHECODE_SE, Phenotype_cateogory$phecode)]
#Map number of distinct se + their phecodes to the phecode cateogory
Phecodefreq=as.data.frame(se_icd %>% select(PHECODE_SE,phecode_category) %>% distinct() %>% group_by(phecode_category)   %>% tally(name = 'number_phecodes'))
se_phecodecat=as.data.frame(se_icd %>% select(medra_se,phecode_category) %>% distinct() %>% group_by(phecode_category) %>% tally(name = 'number_se'))
Phefre=merge(Phecodefreq, se_phecodecat,by='phecode_category')
Phefre1=melt(Phefre,id.var='phecode_category')                         

ggplot(Phefre1, aes(x=phecode_category, y=value, fill=variable))+ geom_bar(stat = "identity",position='dodge') +theme_classic()+theme(axis.text.x = element_text(angle=-90)) + labs(x='Phecode cateogories', y='Number of side effects') 

```

### Match to entrez id  

```{r, echo=F}

drugbank=fread('/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/Target Drug-UniProt Links_ALL.csv',data.table=F, select=c(2,4))
#drugs and uniprot id
drugbank$Name=str_to_lower(drugbank$Name)
colnames(drugbank)[1]='drugname'
GeneDic <- fread("/Users/aineduffy/Documents/PhD/DrugGenetics/Data_downloaded/GeneDictionary.txt", data.table = FALSE, select=c(2,19,26))
drugbank$entrez_id=GeneDic$entrez_id[match(drugbank$`UniProt ID`, GeneDic$uniprot_ids)]
drugbank$symbol=GeneDic$symbol[match(drugbank$`UniProt ID`, GeneDic$uniprot_ids)]
drugbank1=drugbank[!is.na(drugbank$entrez_id),]


se_icd_gene=inner_join(se_icd, drugbank[-2], by='drugname') #file -412794 rows
se_icd_gene=se_icd_gene[!is.na(se_icd_gene$entrez_id),]
# #664 drugs 

print(paste('There are',length(unique(se_icd_gene$drugname)), 'unique drugs'))
print(paste('There are',length(unique(se_icd_gene$entrez_id)), 'unique genes'))
print(paste('There are',length(unique(se_icd_gene$medra_se)), 'unique se and out of these', length(unique(se_icd_gene$medra_se[!is.na(se_icd_gene$se_adr)])), 'has an adr score'))
print(paste('There are',length(unique(se_icd_gene$PHECODE_SE)), 'unique phecodes and out of these', length(unique(se_icd_gene$PHECODE_SE[!is.na(se_icd_gene$se_adr)])), 'has an adr score'))


####run with sider 4. Only did sider 2 becuase no adr score. now if imputing them can do with s4
```

